### 2024-10-21 paper/paper.qmd
add introduction and variable overview in paper

---
title: "Forecasting the US Presidential Election: A Poll-of-Polls Approach Using Linear Models"
author: 
  - Ziqi Zhu
  - Yuanchen Miao
  - author3
thanks: "Code and data are available at: https://github.com/zzq20010617/2024_USelection_prediction"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

This paper examines the development of a multiple linear regression (MLR) model to predict the percentage of support (pct) for U.S. presidential candidates based on polling data. The data includes a range of predictors, such as sample size, pollster quality factors, and time-related factors. The focus of the analysis is on nationwide polling data, aggregated from different pollsters, to create a robust model for forecasting election outcomes. The goal is to provide a clearer understanding of how various poll attributes influence polling results and to derive insights that can predict election outcomes.

The primary estimand in this study is the expected percentage of support (pct) for a U.S. presidential candidate, given polling data and relevant predictor variables like rating of the pollster and sample size of the poll.

Our Results shows that

This analysis is useful in understanding how various polling factors contribute to predicting election results. By modeling the relationship between polling result and factors, the study enhances the ability to forecast election outcomes based on public opinion data. This model provides a practical tool for researchers, political strategists, and analysts to assess the reliability of polling data and its implications for elections.

# Data {#sec-data}

## Overview

Our data is download from [@polls], the website gathered survey data from different pollsters, We use the statistical programming language R [@citeR], and packages [@lubridate], [@dplyr], [@tidyverse], to process the data. Following @tellingstories, we consider...

## Measurement

The dataset is about public opinions about U.S. presidential candidates, which are captured through polling.

## Outcome variables

Our primary outcome variable is **support percentage (pct)**, which represents the percentage of respondents who support each candidate. This variable is modeled as the response variable in the MLR analyses.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

# ggplot(penguins, aes(x = island, fill = species)) +
#   geom_bar(alpha = 0.8) +
#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
#                     guide = "none") +
#   theme_minimal() +
#   facet_wrap(~species, ncol = 1) +
#   coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

### Sample Size

Sample size is an important predictor because it influences the reliability of a poll. Larger sample sizes tend to reduce sampling error, providing more accurate reflections of voter sentiment. In our MLR model, the sample size is log-transformed to account for diminishing returns—larger polls do not necessarily offer proportionally better accuracy.

### Poll Score

The poll score is a measure of the error and bias we can attribute to a pollster, negative number is better.

### Numeric Grade

This variable is an aggregate score of the poll based on a numeric scale. It serves as an indicator of the pollster's historical performance and the methodology used. 3 is the maximum and some pollsters have no rating.

### Transparency Score

Transparency score measures how openly a pollster reports their methodology and results. A higher transparency score suggests that the pollster has disclosed key details about how the poll was conducted, improving trust in the poll results.

### Days Since Start

This variable represents the number of days since the beginning of the election polling period. It helps capture the dynamic nature of voter preferences over time, accounting for shifts in public opinion as the election date nears.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-21 scripts/06-model_data.R

@@ -25,18 +25,18 @@ nationwide_data <- nationwide_data %>%
   mutate(days_since_start = as.numeric(end_date - min(end_date)))
 
 unique_parties <- unique(nationwide_data$party)
-
-# model that include party as a predictor
-# model <- lm(pct ~ log(sample_size) + pollscore + numeric_grade + transparency_score+days_since_start+party, data = nationwide_data)
-# plot(model, which = 1)
-# new_data <- data.frame(
-#   sample_size = 1500,
-#   pollscore = -1.5,
-#   numeric_grade = 3,
-#   transparency_score = 8,
-#   days_since_start = 310,
-#   party = "DEM"
-# )
+nationwide_data$party <- as.factor(nationwide_data$party)
+#model that include party as a predictor
+model <- lm(pct ~ log(sample_size) + pollscore + numeric_grade + transparency_score+days_since_start+party, data = nationwide_data)
+plot(model, which = 1)
+new_data <- data.frame(
+  sample_size = 1500,
+  pollscore = -1.5,
+  numeric_grade = 3,
+  transparency_score = 8,
+  days_since_start = 310,
+  party = factor("DEM", levels = levels(nationwide_data$party))
+)
 
 
 ### Model data ####

### 2024-10-21 scripts/example.R

@@ -1,0 +1,228 @@
+#### Set-up ####
+# Load libraries
+library(tidyverse)
+library(janitor)
+library(lubridate)
+library(broom)
+library(modelsummary)
+library(rstanarm)
+library(splines)
+
+
+#### Prepare dataset ####
+# Read in the data and clean variable names
+data <- read_csv("data/01-raw_data/raw_data.csv") |>
+  clean_names()
+
+# Filter data to Harris estimates based on high-quality polls after she declared
+just_harris_high_quality <- data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
+    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+  )
+
+
+#### Plot data ####
+base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+# Plots poll estimates and overall smoothing
+base_plot +
+  geom_point() +
+  geom_smooth()
+
+# Color by pollster
+# This gets messy - need to add a filter - see line 21
+base_plot +
+  geom_point(aes(color = pollster)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+# Facet by pollster
+# Make the line 21 issue obvious
+# Also - is there duplication???? Need to go back and check
+base_plot +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(vars(pollster))
+
+# Color by pollscore
+base_plot +
+  geom_point(aes(color = factor(pollscore))) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+
+#### Starter models ####
+# Model 1: pct as a function of end_date
+model_date <- lm(pct ~ end_date, data = just_harris_high_quality)
+
+# Model 2: pct as a function of end_date and pollster
+model_date_pollster <- lm(pct ~ end_date + pollster, data = just_harris_high_quality)
+
[…]

### 2024-10-21 paper.qmd
Pollster in appendix is updated

--- 2024-10-21 paper/paper.qmd
@@
   - Yuanchen Miao
-  - author3
+  - Claire Chang
 thanks: "Code and data are available at: https://github.com/zzq20010617/2024_USelection_prediction"
@@
 
+## Trafalgar Group’s Methodology
+
+The Trafalgar Group conducts polls ranging from major political campaigns to marketing surveys.
+The organization's 0.7 pollster rating indicates moderate accuracy and reliability compared to other pollsters.
+The population consists of all eligible voters in the U.S., while the sampling frame likely includes registered voters across states, segmented by demographics like age, gender, and party affiliation.
+Trafalgar Group typically samples likely voters, focusing on those most likely to participate in upcoming elections.
+Trafalgar Group recruits its sample using a mix of interactive voice response (IVR), live phone calls, text messages, online panels, and email surveys.
+Trafalgar Group employs stratified random sampling to ensure proportional representation of subgroups like political party, gender, and region.
+This method improves accuracy by reflecting the electorate's demographic and political makeup but may risk over-stratification, giving smaller voter groups disproportionate influence and potentially skewing results.
+Trafalgar Group handles non-response by using weighting to adjust the sample, ensuring respondent demographics (age, race, education) match population parameters.
+This method reduces bias by accounting for underrepresented groups, though heavy reliance on post-survey adjustments may introduce new biases, particularly when there are large discrepancies in response rates across demographics.
+
 \newpage

### 2024-10-21 paper/paper.qmd
Name and appendix is updated with pollster

@@
 
-# Additional data details
+# Trafalgar Group’s Methodology
+
+The Trafalgar Group conducts polls ranging from major political campaigns to marketing surveys.
+The organization's 0.7 pollster rating indicates moderate accuracy and reliability compared to other pollsters.
+The population consists of all eligible voters in the U.S., while the sampling frame likely includes registered voters across states, segmented by demographics like age, gender, and party affiliation.
+Trafalgar Group typically samples likely voters, focusing on those most likely to participate in upcoming elections.
+Trafalgar Group recruits its sample using a mix of interactive voice response (IVR), live phone calls, text messages, online panels, and email surveys.
+Trafalgar Group employs stratified random sampling to ensure proportional representation of subgroups like political party, gender, and region.
+This method improves accuracy by reflecting the electorate's demographic and political makeup but may risk over-stratification, giving smaller voter groups disproportionate influence and potentially skewing results.
+Trafalgar Group handles non-response by using weighting to adjust the sample, ensuring respondent demographics (age, race, education) match population parameters.
+This method reduces bias by accounting for underrepresented groups, though heavy reliance on post-survey adjustments may introduce new biases, particularly when there are large discrepancies in response rates across demographics.
 

### 2024-10-21 paper/paper.qmd
Appendix updated with pollster

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "1) what was done, 2) what was found, and 3) why this matters. 4) sentence."
 number-sections: true
@@
 
-# Trafalgar Group’s Methodology
+# Idealized methodology
+
+# Idealized Survey
+
+# Trafalgar group’s methodology overview and evaluation
 
@@
 The organization's 0.7 pollster rating indicates moderate accuracy and reliability compared to other pollsters.
-The population consists of all eligible voters in the U.S., while the sampling frame likely includes registered voters across states, segmented by demographics like age, gender, and party affiliation.
+The population consists of all eligible voters in the U.S., while the sampling frame includes registered voters across states, segmented by demographics like age, gender, party affiliation, and ethnicity.
 Trafalgar Group typically samples likely voters, focusing on those most likely to participate in upcoming elections.
@@
 This method improves accuracy by reflecting the electorate's demographic and political makeup but may risk over-stratification, giving smaller voter groups disproportionate influence and potentially skewing results.
-Trafalgar Group handles non-response by using weighting to adjust the sample, ensuring respondent demographics (age, race, education) match population parameters.
+Trafalgar Group handles non-response by using weighting to adjust the sample, ensuring respondent demographics match population parameters.
 This method reduces bias by accounting for underrepresented groups, though heavy reliance on post-survey adjustments may introduce new biases, particularly when there are large discrepancies in response rates across demographics.

### 2024-10-21 scripts/03-clean_data.R
update clean data script

@@ -28,14 +28,11 @@ cleaned_data <-
   raw_data |>
   janitor::clean_names() %>%
   mutate(end_date = mdy(end_date)) %>%
+  mutate(
+    state = if_else(is.na(state), "National", state)) %>%
   filter(year(end_date) == 2024)|>
   select(pollster,sample_size,numeric_grade,pollscore,state,transparency_score,end_date,party,answer,pct)
 
 
-# filter by state to keep those polls nationwide
-nationwide_data <- cleaned_data %>%
-  filter(is.na(state)) |>
-  select(-state)
-
 #### Save data ####
-write_csv(nationwide_data, "data/02-analysis_data/nationwide_data.csv")
+write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")

### 2024-10-21 scripts/00-simulate_data.R
simulated data and tests

@@ -17,18 +17,70 @@ set.seed(853)
 #### Simulate data ####
 # State names
 states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+  "Alabama",
+  "Alaska",
+  "Arizona",
+  "Arkansas",
+  "California",
+  "Colorado",
+  "Connecticut",
+  "Delaware",
+  "Florida",
+  "Georgia",
+  "Hawaii",
+  "Idaho",
+  "Illinois",
+  "Indiana",
+  "Lowa",
+  "Kansas",
+  "Kentucky",
+  "Louisiana",
+  "Maine",
+  "Maryland",
+  "Massachusetts",
+  "Michigan",
+  "Minnesota",
+  "Mississippi",
+  "Missouri",
+  "Montana",
+  "Nebraska",
+  "Nevada",
+  "New Hampshire",
+  "New Jersey",
+  "New Mexico",
+  "New York",
+  "North Carolina",
+  "North Dakota",
+  "Ohio",
+  "Oklahoma",
+  "Oregon",
+  "Pennsylvania",
+  "Rhode Island",
+  "South Carolina",
+  "South Dakota",
+  "Tennessee",
+  "Texas",
+  "Utah",
+  "Vermont",
+  "Virginia",
+  "Washington",
+  "West Virginia",
+  "Wisconsin",
+  "Wyoming"
 )
 
 # Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+parties <- c("Republican", "Democratic", "Greens", "Libertarians", "Constitution", "Natural Law")
+
+# Initialize a vector of length 50
+extended_prob <- numeric(50)
+
+# Assign the original probabilities to the first 8 elements
+prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025)
+extended_prob[1:8] <- prob
+
+# Distribute the remaining probability mass (which is 0 in this case) equally among the rest of the elements
+extended_prob[9:50] <- (1 - sum(extended_prob)) / 42
 
 # Create a dataset by randomly assigning states and parties to divisions
 analysis_data <- tibble(
@@ -37,14 +89,15 @@ analysis_data <- tibble(
     states,
     size = 151,
     replace = TRUE,
[…]

### 2024-10-21 scripts/01-test_simulated_data.R

@@ -34,10 +34,10 @@ if (nrow(analysis_data) == 151) {
 }
 
 # Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+if (ncol(analysis_data) == 4) {
+  message("Test Passed: The dataset has 4 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 4 columns.")
 }
 
 # Check if all values in the 'division' column are unique
@@ -48,18 +48,67 @@ if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
 }
 
 # Check if the 'state' column contains only valid Australian state names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
-                  "Western Australia", "Tasmania", "Northern Territory", 
-                  "Australian Capital Territory")
+valid_states <- c(
+  "Alabama",
+  "Alaska",
+  "Arizona",
+  "Arkansas",
+  "California",
+  "Colorado",
+  "Connecticut",
+  "Delaware",
+  "Florida",
+  "Georgia",
+  "Hawaii",
+  "Idaho",
+  "Illinois",
+  "Indiana",
+  "Lowa",
+  "Kansas",
+  "Kentucky",
+  "Louisiana",
+  "Maine",
+  "Maryland",
+  "Massachusetts",
+  "Michigan",
+  "Minnesota",
+  "Mississippi",
+  "Missouri",
+  "Montana",
+  "Nebraska",
+  "Nevada",
+  "New Hampshire",
+  "New Jersey",
+  "New Mexico",
+  "New York",
+  "North Carolina",
+  "North Dakota",
+  "Ohio",
+  "Oklahoma",
+  "Oregon",
+  "Pennsylvania",
+  "Rhode Island",
+  "South Carolina",
+  "South Dakota",
+  "Tennessee",
+  "Texas",
+  "Utah",
+  "Vermont",
+  "Virginia",
+  "Washington",
+  "West Virginia",
+  "Wisconsin",
+  "Wyoming"
+)
 
 if (all(analysis_data$state %in% valid_states)) {
-  message("Test Passed: The 'state' column contains only valid Australian state names.")
+  message("Test Passed: The 'state' column contains only valid United states state names.")
 } else {
   stop("Test Failed: The 'state' column contains invalid state names.")
 }
 
 # Check if the 'party' column contains only valid party names
[…]

### 2024-10-22 paper/paper.qmd
add model scripts, modify paper

@@
 Our primary outcome variable is **support percentage (pct)**, which represents the percentage of respondents who support each candidate.
-This variable is modeled as the response variable in the MLR analyses.
+This variable is modeled as the response variable in the MLR analyse.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Our model is build from a transformed version of the MLR in above section, with transparency_score dropped as it has a significant level greater than 0.05.
+
+$y_i$ represent the number of individuals in the sample that support the Democratic party (this corresponds to num_party).
+The response is modeled as binomial: $y_i \sim \text{Binomial}(\text{sample\_size}_i, p_i)$, where $p_i$ is the probability that an individual in poll $i$ supports the Democratic party, and $\text{sample\_size}_i$ is the total number of individuals surveyed in poll $i$.
 
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i | p_i &\sim \mbox{Binomial}(\text{sample\_size}_i, p_i) \\
+\log \left( \frac{p_i}{1 - p_i} \right) &= \beta_0 + \beta_1 \log(\text{sample\_size}_i) + \beta_2 \, \text{pollscore}_i + \beta_3 \, \text{numeric\_grade}_i + \alpha_j \\
+\alpha_j &\sim \mbox{Normal}(0, \sigma_{\text{pollster}}) \\
+\beta_0 &\sim \mbox{Normal}(0, 5) \\
+\beta_1 &\sim \mbox{Normal}(0, 5) \\
+\beta_2 &\sim \mbox{Normal}(0, 5) \\
+\beta_3 &\sim \mbox{Normal}(0, 5) \\
+\sigma_{\text{pollster}} &\sim \mbox{Exponential}(1)
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We first priors as default with “autoscale = TRUE”,
 

### 2024-10-22 scripts/06-model_data.R

@@ -13,7 +13,8 @@ library(tidyverse)
 #library(rstanarm)
 
 #### Read data ####
-nationwide_data <- read_csv("data/02-analysis_data/nationwide_data.csv")
+cleaned_data <- read_csv("data/02-analysis_data/cleaned_data.csv")
+nationwide_data <- cleaned_data %>%  filter(state == "National")
 
 # only two entries for Conservative Party
 unique_parties <- unique(nationwide_data$party)
@@ -26,18 +27,6 @@ nationwide_data <- nationwide_data %>%
 
 unique_parties <- unique(nationwide_data$party)
 nationwide_data$party <- as.factor(nationwide_data$party)
-#model that include party as a predictor
-model <- lm(pct ~ log(sample_size) + pollscore + numeric_grade + transparency_score+days_since_start+party, data = nationwide_data)
-plot(model, which = 1)
-new_data <- data.frame(
-  sample_size = 1500,
-  pollscore = -1.5,
-  numeric_grade = 3,
-  transparency_score = 8,
-  days_since_start = 310,
-  party = factor("DEM", levels = levels(nationwide_data$party))
-)
-
 
 ### Model data ####
 boxplot(pct ~ party, data = nationwide_data)
@@ -68,19 +57,189 @@ for (party in names(models)) {
 }
 
 # choose a party to check summary
-model_summary <- summary(models[['DEM']])
-
-# Print the model summary
-print(model_summary)
-
+model_summary <- summary(models[['REP']])
 # Access the R-squared and Adjusted R-squared values
 r_squared <- model_summary$r.squared
 adjusted_r_squared <- model_summary$adj.r.squared
 
+# Print the model summary
+print(model_summary)
+
+# Need number not percent for some models
+nationwide_data <- nationwide_data %>% mutate(
+  num_party = round((pct / 100) * sample_size, 0)
+)
+# filter nationwide_data by parties
+dem_data <- nationwide_data %>%
+  filter(party == "DEM")
+dem_data <- na.omit(dem_data)
+
+dem_data <- dem_data |>
+  mutate(
+    pollster = factor(pollster)
+  )
+
+rep_data <- nationwide_data %>%
+  filter(party == "REP")
+rep_data <- na.omit(rep_data)
+
+rep_data <- rep_data |>
+  mutate(
+    pollster = factor(pollster)
+  )
+
+
+# Augment the datasets with predictions
+dem_data <- dem_data |>
+  mutate(fitted_pct_dem = predict(models[['DEM']]))
[…]

### 2024-10-22 scripts/example.R

@@ -152,7 +152,7 @@ plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
 plot(bayesian_model_2, pars = "(Intercept)", prob = 0.95)
 
 # Model summary works the same as above for Bayesian models.
-
+# modelsummary(models = list("Model 1" = bayesian_model_1, "Model 2" = bayesian_model_2))
 
 #### Bayesian models and splines ####
 # Change date to be number of days since she declared - it's a counter not a date

### 2024-10-22 paper/paper.qmd
Abstract updated and format changed in appendix

@@
 date-format: long
-abstract: "1) what was done, 2) what was found, and 3) why this matters. 4) sentence."
+abstract: "This paper presents a multiple linear regression model for forecasting the outcome of the 2024 U.S.
+Presidential Election.
+Using nationwide polling data, we predict the percentage of support for U.S.
+presidential candidates based on key factors such as sample size, pollster quality, and timing.
+Our model aggregates these predictions to simulate potential election outcomes.
+The results reveal how specific poll attributes affect the accuracy of forecasts, offering valuable insights for predicting the likelihood of various electoral scenarios."
 number-sections: true
@@
 
-# Idealized methodology
+## Idealized  Survey and Methodology
 
-# Idealized Survey
-
-# Trafalgar group’s methodology overview and evaluation
+## Trafalgar group’s methodology overview and evaluation
 

### 2024-10-22 scripts/03-clean_data.R
the way of how to filter the date after 2024 is changed

@@ -21,18 +21,21 @@ library(dplyr)
 #Party Affiliation: Voters' support tends to correlate with their party.
 #Pollster Rating: Reliability and bias rating of the pollster.
 
+# Read and load raw data
 raw_data <- read_csv("data/01-raw_data/raw_data.csv")
+
+# change the type of end date in raw data from double to date
+raw_data$end_date <- as.character(raw_data$end_date)
+raw_data$end_date <- mdy(raw_data$end_date)
 
 # select useful columns and filter by date to keep polls ends in 2024
 cleaned_data <-
   raw_data |>
   janitor::clean_names() %>%
-  mutate(end_date = mdy(end_date)) %>%
+  select(pollster,sample_size,numeric_grade,pollscore,state,transparency_score,end_date,party,answer,pct) %>%
+  filter(end_date > as.Date("2024-01-01")) %>%
   mutate(
-    state = if_else(is.na(state), "National", state)) %>%
-  filter(year(end_date) == 2024)|>
-  select(pollster,sample_size,numeric_grade,pollscore,state,transparency_score,end_date,party,answer,pct)
-
+    state = if_else(is.na(state), "National", state))
 
 #### Save data ####
 write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")

### 2024-10-22 paper/paper.qmd
data section of paper and reference is updated

--- 2024-10-22 paper/paper.qmd
@@
 
-Our data is download from [@polls], the website gathered survey data from different pollsters, We use the statistical programming language R [@citeR], and packages [@lubridate], [@dplyr], [@tidyverse], to process the data.
-Following @tellingstories, we consider...
+Our data is download from [Five Thirty Eight](https://projects.fivethirtyeight.com/polls)[@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate**[@lubridate], **dplyr**[@dplyr], **tidyverse**[@tidyverse], **model summary**[@models] to process the data.
+Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
+Graphs are created by using ggplot2 in from package **tidyverse**[@tidyverse]Following **Telling stories with data**[@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
+We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
 

### 2024-10-22 scripts/04-test_analysis_data.R
analysis data is tested

@@ -12,58 +12,124 @@
 library(tidyverse)
 library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analisis_data/cleaned_data.csv")
 
+# Test if the data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
+# Check if the dataset has 151 rows
+if (nrow(analysis_data) == 151) {
+  message("Test Passed: The dataset has 151 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 151 rows.")
+}
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
+# Check if the dataset has 3 columns
+if (ncol(analysis_data) == 4) {
+  message("Test Passed: The dataset has 4 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 4 columns.")
+}
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
+# Check if all values in the 'division' column are unique
+if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
+  message("Test Passed: All values in 'division' are unique.")
+} else {
+  stop("Test Failed: The 'division' column contains duplicate values.")
+}
 
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
-})
+# Check if the 'state' column contains only valid Australian state names
+valid_states <- c(
+  "Alabama",
+  "Alaska",
+  "Arizona",
+  "Arkansas",
+  "California",
+  "Colorado",
+  "Connecticut",
+  "Delaware",
+  "Florida",
+  "Georgia",
+  "Hawaii",
+  "Idaho",
+  "Illinois",
+  "Indiana",
+  "Lowa",
+  "Kansas",
[…]

### 2024-10-22 scripts/01-test_simulated_data.R
fix minor error of simulate and cleaned data test

@@ -33,7 +33,7 @@ if (nrow(analysis_data) == 151) {
   stop("Test Failed: The dataset does not have 151 rows.")
 }
 
-# Check if the dataset has 3 columns
+# Check if the dataset has 4 columns
 if (ncol(analysis_data) == 4) {
   message("Test Passed: The dataset has 4 columns.")
 } else {
@@ -47,7 +47,7 @@ if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
   stop("Test Failed: The 'division' column contains duplicate values.")
 }
 
-# Check if the 'state' column contains only valid Australian state names
+# Check if the 'state' column contains only valid United state names
 valid_states <- c(
   "Alabama",
   "Alaska",

### 2024-10-22 paper/paper.qmd
title changed

@@
 ---
-title: "Forecasting the US Presidential Election: A Poll-of-Polls Approach Using Linear Models"
+title: "Forecasting the US Presidential Election: A Poll-of-Polls Approach Using Linear Models By Using Data of Pollsters From Five Thirty Eight In 2024"
 author: 

### 2024-10-22 scripts/03-clean_data.R
add a harris only cleaned_data

@@ -37,5 +37,27 @@ cleaned_data <-
   mutate(
     state = if_else(is.na(state), "National", state))
 
+
+#### Prepare dataset for Harris only ####
+# Read in the data and clean variable names
+data <- read_csv("data/01-raw_data/raw_data.csv") |>
+  janitor::clean_names()
+
+# Filter data to Harris estimates based on high-quality polls after she declared
+just_harris_high_quality <- data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= mean(!is.na(numeric_grade))
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> 
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0) 
+  )
+
 #### Save data ####
 write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")
+write_csv(just_harris_high_quality, "data/02-analysis_data/just_harris_high_quality.csv")

### 2024-10-22 scripts/06-model_data.R
add a harris only model

@@ -235,6 +235,27 @@ ggplot(combined_data, aes(x = days_since_start, y = pct, color = pollster)) +
   ) +
   theme_minimal()
 
+#### Just Harris model ####
+
+# Model 1: pct as a function of end_date
+model_date <- lm(pct ~ end_date, data = just_harris_high_quality)
+
+# Augment data with model predictions
+just_harris_high_quality <- just_harris_high_quality |>
+  mutate(
+    fitted_date = predict(model_date),
+  )
+
+# Plot model predictions
+# Model 1
+ggplot(just_harris_high_quality, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_date), color = "blue", linetype = "dotted") +
+  theme_classic() +
+  scale_y_continuous(labels = scales::percent_format(scale = 1))+
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
+
+
 
 #### Save model ####
 saveRDS(

### 2024-10-22 scripts/04-test_cleaned_data.R
fix minor error in test_cleaned_data

@@ -1,0 +1,133 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the simulated Australian 
+#electoral divisions dataset.
+# Author: Rohan Alexander
+# Date: 26 September 2024
+# Contact: rohan.alexander@utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - 00-simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `starter_folder` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+# Test if the data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+
+#### Test data ####
+
+
+
+# Check if the dataset has 4 columns
+if (ncol(analysis_data) == 4) {
+  message("Test Passed: The dataset has 4 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 4 columns.")
+}
+
+# Check if all values in the 'division' column are unique
+if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
+  message("Test Passed: All values in 'division' are unique.")
+} else {
+  stop("Test Failed: The 'division' column contains duplicate values.")
+}
+
+# Check if the 'state' column contains only valid United state names
+valid_states <- c(
+  "Alabama",
+  "Alaska",
+  "Arizona",
+  "Arkansas",
+  "California",
+  "Colorado",
+  "Connecticut",
+  "Delaware",
+  "Florida",
+  "Georgia",
+  "Hawaii",
+  "Idaho",
+  "Illinois",
+  "Indiana",
+  "Lowa",
+  "Kansas",
+  "Kentucky",
+  "Louisiana",
+  "Maine",
+  "Maryland",
+  "Massachusetts",
+  "Michigan",
+  "Minnesota",
+  "Mississippi",
+  "Missouri",
+  "Montana",
+  "Nebraska",
+  "Nevada",
+  "New Hampshire",
+  "New Jersey",
+  "New Mexico",
+  "New York",
+  "North Carolina",
+  "North Dakota",
+  "Ohio",
+  "Oklahoma",
+  "Oregon",
+  "Pennsylvania",
[…]

### 2024-10-22 paper/paper.qmd
save spline fit as models, update paper in model section

--- 2024-10-22 paper/paper.qmd
@@
 
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+cleaned_data <- read_csv(here::here("data/02-analysis_data/cleaned_data.csv"))
+nationwide_data <- cleaned_data %>%  filter(state == "National")
+
+# only two entries for Conservative Party
+unique_parties <- unique(nationwide_data$party)
+
+# drop columns that have NA in numeric_grade or pollscore, and 2 entries for CON party
+nationwide_data <- nationwide_data %>%
+  drop_na(numeric_grade, pollscore) %>%
+  filter(party != "CON") %>%
+  mutate(days_since_start = as.numeric(end_date - min(end_date)))
+
+unique_parties <- unique(nationwide_data$party)
+nationwide_data$party <- as.factor(nationwide_data$party)
+
+nationwide_data <- nationwide_data %>% mutate(
+  num_party = round((pct / 100) * sample_size, 0)
+)
+# filter nationwide_data by parties
+dem_data <- nationwide_data %>%
+  filter(party == "DEM")
+dem_data <- na.omit(dem_data)
+
+dem_data <- dem_data |>
+  mutate(
+    pollster = factor(pollster)
+  )
+
+rep_data <- nationwide_data %>%
+  filter(party == "REP")
+rep_data <- na.omit(rep_data)
+
+rep_data <- rep_data |>
+  mutate(
+    pollster = factor(pollster)
+  )
+```
+
 # Introduction
@@
 
-Our data is download from [Five Thirty Eight](https://projects.fivethirtyeight.com/polls)[@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate**[@lubridate], **dplyr**[@dplyr], **tidyverse**[@tidyverse], **model summary**[@models] to process the data.
-Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
-Graphs are created by using ggplot2 in from package **tidyverse**[@tidyverse]Following **Telling stories with data**[@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
-We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
+Our data is download from [@polls], the website gathered survey data from different pollsters, We use the statistical programming language R [@citeR], and packages [@lubridate], [@dplyr], [@tidyverse], to process the data.
+Following @tellingstories, we consider...
 
@@
 
-## Outcome variables
+## Outcome variables {#sec-octvar}
 
@@
 
-## Predictor variables
+## Predictor variables {#sec-prevar}
 
@@
 
-# Model
+# Models
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modelling strategy is try to capture the. Firstly,...
+In the following section we briefly describe the models we used to investigate. Background details and diagnostics are included in [Appendix -@sec-model-details].
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+## SLR Model set-up
 
-## Model set-up
-
-Our model is build from a transformed version of the MLR in above section, with transparency_score dropped as it has a significant level greater than 0.05.
-
-$y_i$ represent the number of individuals in the sample that support the Democratic party (this corresponds to num_party).
-The response is modeled as binomial: $y_i \sim \text{Binomial}(\text{sample\_size}_i, p_i)$, where $p_i$ is the probability that an individual in poll $i$ supports the Democratic party, and $\text{sample\_size}_i$ is the total number of individuals surveyed in poll $i$.
-
+## MLR Model set-up
+We first filtered our data by different parties, as the single linear model but include other 3 parties(GRE, IND, and LIB, Conservative only has two entries in cleaned nationwide data and has 0 pct so its been removed), and fit linear model for each of them.
+The formula for Multiple linear regression is as follow
 \begin{align}
-y_i | p_i &\sim \mbox{Binomial}(\text{sample\_size}_i, p_i) \\
-\log \left( \frac{p_i}{1 - p_i} \right) &= \beta_0 + \beta_1 \log(\text{sample\_size}_i) + \beta_2 \, \text{pollscore}_i + \beta_3 \, \text{numeric\_grade}_i + \alpha_j \\
-\alpha_j &\sim \mbox{Normal}(0, \sigma_{\text{pollster}}) \\
-\beta_0 &\sim \mbox{Normal}(0, 5) \\
-\beta_1 &\sim \mbox{Normal}(0, 5) \\
-\beta_2 &\sim \mbox{Normal}(0, 5) \\
-\beta_3 &\sim \mbox{Normal}(0, 5) \\
-\sigma_{\text{pollster}} &\sim \mbox{Exponential}(1)
+\text{pct}_i &= \beta_0 + \beta_1 \log(\text{sample\_size}_i) + \beta_2 \, \text{pollscore}_i + \beta_3 \, \text{numeric\_grade}_i \\
+&+ \beta_4 \, \text{transparency\_score}_i + \beta_5 \, \text{days\_since\_start}_i + \epsilon_i
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We first priors as default with “autoscale = TRUE”,
+Response Variable is pct. Predictors are log(sample_size), pollscore, numeric_grade, transparency_score, and days_since_start, detail can be found in [@sec-octvar], and [@sec-prevar]
 
-### Model justification
+### Result
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+Our results for the MLR model are summarized in @tbl-mlrmodelresults.
 
@@
 
-library(rstanarm)
-
-# first_model <-
-#   readRDS(file = here::here("models/first_model.rds"))
+MLR_models <-
+  readRDS(file = here::here("models/MLRmodels.rds"))
 ```
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-mlrmodelresults
+#| tbl-cap: "Explanatory models of pct based on pollscore, sample size, and time"
 #| warning: false
 
-# modelsummary::modelsummary(
-#   list(
-#     "First model" = first_model
-#   ),
-#   statistic = "mad",
-#   fmt = 2
-# )
+modelsummary::modelsummary(
+  MLR_models,
+  fmt = 2
+)
 ```
+We first check the residual vs.
+fitted plot as shown in [Appendix @sec-model-details], and see no obvious violation of assumptions.
+According to the summary, DEM and REP models are based on the largest number of observations (both 1235).
+Smaller parties have fewer observations, which may lead to less robust models.
+The $r^2$ shows that the model explain the highest amount of variance on Democratic Party than others with value 0.357, and other parties less than 0.2.
+Based on the summary of models for DEM and REP [@tbl-mlrmodelsum], we are not sure if those predictor that measure quality of polls are significant or not, so we keep them for now in Bayesian model.
+Prediction of MLR models is shown in [@fig-mlrpredict], which we can see blue line which is Democratic party has a steeper slope and start to take the lead between day 100 and day 200.
+
+```{r}
+#| echo: false
+#| label: fig-mlrpredict
+#| fig-cap: "Prediction with MLR models for DEM and REP party"
+#| warning: false
+
+# Augment the datasets with predictions
+dem_data <- dem_data |>
+  mutate(fitted_pct_dem = predict(MLR_models[['DEM']]))
+rep_data <- rep_data |>
+  mutate(fitted_pct_rep = predict(MLR_models[['REP']]))
+
+# Combine both datasets into one
+combined_data <- bind_rows(dem_data, rep_data)
+
+# Plot with colors based on the party
+ggplot(combined_data, aes(x = days_since_start)) +
+  geom_line(aes(y = fitted_pct_dem, color = "DEM")) +
+  geom_line(aes(y = fitted_pct_rep, color = "REP")) +
+  scale_color_manual(values = c("DEM" = "blue", "REP" = "red"), 
+                     name = "Party",
+                     labels = c("DEM (Blue)", "REP (Red)")) +
+  theme_classic() +
+  labs(y = "Percent", x = "Days") +
+  theme(legend.position = "top")
+```
+
+## Bayesian Model set-up
+The Bayesian model is build from the MLR models from above section.
+With the inspiration from the example R code, we introduce the random effects for Pollsters, and we use a logistic link to predict the probability of support, modeling count data which is in binomial distribution.
+We choose to use a default prior and enable autoscale, the formula is shown below
+
+\begin{align} 
+\log \left( \frac{p_i}{1 - p_i} \right) &= \beta_0 + \beta_1 \log(\text{sample\_size}_i) + \beta_2 \, \text{pollscore}_i + \beta_3 \, \text{numeric\_grade}_i \\
+&+ \beta_4 \, \text{transparency\_score}_i + \beta_5 \, \text{days\_since\_start}_i + \alpha_j
+\end{align}
+\alpha_j &\sim \text{Normal}(0, \sigma_{\text{pollster}}) \\
+\beta_0 &\sim \text{Normal}(0, 2.5) \\
+\beta_1 &\sim \text{Normal}(0, 2.5) \\
+\beta_2 &\sim \text{Normal}(0, 2.5) \\
+\beta_3 &\sim \text{Normal}(0, 2.5) \\
+\beta_4 &\sim \text{Normal}(0, 2.5) \\
+\beta_5 &\sim \text{Normal}(0, 2.5) \\
+\sigma_{\text{pollster}} &\sim \text{Exponential}(1)
+\end{align}
+
+$y_i$ represent the number of individuals in the sample that support the Democratic party (this corresponds to num_party).
+The response is modeled as binomial: $y_i \sim \text{Binomial}(\text{sample\_size}_i, p_i)$, where $p_i$ is the probability that an individual in poll $i$ supports the Democratic party, and $\text{sample\_size}_i$ is the total number of individuals surveyed in poll $i$.
+
+### Result
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+Bay_model <-
+  readRDS(file = here::here("models/bayesian_model_dem.rds"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-baymodelresults
+#| tbl-cap: ""
+#| warning: false
+
+modelsummary::modelsummary(
+  Bay_model,
+  fmt = 2
+)
+```
+
+# Prediction
+
+To get a prediction by the Vayesian model we have, we first spline fit the model, and then create a data frame and generate posterior predictions by the new data frame.
+## Spline Fit for Bayesian model set-up
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm.
+We use priors $\text{Normal}(0, 5) \\$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
+Then we crate a data frame like [@tbl-newdata], we set the quality scale of predict data to the best to perform a poll that has high quality.
+the result is shown as [@sec-splresult]
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-newdata
+
+# Create new data for prediction
+new_data <- data.frame(
+  pollscore = -1.5,
+  sample_size = 1200,
+  numeric_grade = 3,
+  transparency_score=10,
+  days_since_start = seq(
+    min(dem_data$days_since_start),
+    max(dem_data$days_since_start),
+    length.out = 100
+  ),
+  pollster = factor("TIPP", levels = levels(dem_data$pollster))
+)
+
+head(new_data)
+```
+
+## Result {#sec-splresult}
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+spline_dem <-
+  readRDS(file = here::here("models/spline_dem.rds"))
+spline_rep <-
+  readRDS(file = here::here("models/spline_rep.rds"))
+```
+
+```{r}
+#| echo: false
+#| label: fig-splpredict
+#| fig-cap: "Poll Percentage over Time with Spline Fit for DEM and REP party"
+#| warning: false
+
+posterior_preds_dem <- posterior_predict(spline_model_dem, newdata = new_data)
+posterior_preds_rep <- posterior_predict(spline_model_rep, newdata = new_data)
+
+combined_data <- bind_rows(dem_data, rep_data)
+
+# Plot the spline fit
+ggplot(combined_data, aes(x = days_since_start, y = pct, color = pollster)) +
+  # DEM predictions
+  geom_line(
+    data = pred_summary %>% filter(party == "DEM"),
+    aes(x = days_since_start, y = pred_mean_dem, color = "DEM"),  # Move color inside aes()
+    inherit.aes = FALSE
+  ) +
+  geom_ribbon(
+    data = pred_summary %>% filter(party == "DEM"),
+    aes(x = days_since_start, ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"),  # Move fill inside aes()
+    alpha = 0.2,
+    inherit.aes = FALSE
+  ) +
+  # REP predictions
+  geom_line(
+    data = pred_summary %>% filter(party == "REP"),
+    aes(x = days_since_start, y = pred_mean_rep, color = "REP"),  # Move color inside aes()
+    inherit.aes = FALSE
+  ) +
+  geom_ribbon(
+    data = pred_summary %>% filter(party == "REP"),
+    aes(x = days_since_start, ymin = pred_lower_rep, ymax = pred_upper_rep, fill = "REP"),  # Move fill inside aes()
+    alpha = 0.2,
+    inherit.aes = FALSE
+  ) +
+  # Use scale_color_manual and scale_fill_manual for colors and fills
+  scale_color_manual(values = c("DEM" = "blue", "REP" = "red"), 
+                     name = "Party",
+                     labels = c("DEM (Blue)", "REP (Red)")) +
+  scale_fill_manual(values = c("DEM" = "blue", "REP" = "red"), 
+                    name = "Party",
+                    labels = c("DEM (Blue)", "REP (Red)")) +
+  labs(
+    x = "Days since 2024",
+    y = "Percentage",
+  ) +
+  theme_minimal()
+```
+The prediction graph contains prediction of spline fit model for two parties, blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
+We can tell that the predict support of DEM party is taking the lead by aroud 2 percent until the last day of the dataset.
 
@@
 
-## First discussion point {#sec-first-point}
+## Crossing point in the prediction  {#sec-first-point}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
+From the prediction of spline fit model [@fig-splpredict], There is a clear increase in the support for the DEM party starting around day 200, while the REP party's support remains more stable with only a slight increase over time.
+The increase in support for the DEM party in the predictions could be attributed to Harris's rise as the Democratic candidate after Biden's exit from the race On July 21, 2024, which is about 230 days from start of 2024.
 
@@
 # Model details {#sec-model-details}
+## Assumption check for MLR models
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-assumptioncheck
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: [""]
+
+for (party in names(MLR_models)) {
+  # Extract the model for the current party
+  current_model <- MLR_models[[party]]
+
+  # Create a Residuals vs Fitted plot for the current model
+  plot(current_model, which = 1, main = paste("Residuals vs Fitted for", party, "Party"))
+
+}
+```
+
+## Significant check for MLR models
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-mlrmodelsum
+#| tbl-cap: "Summary of DEM and REP data fit by MLR model"
+#| warning: false
+
+summary(MLR_models[['DEM']])
+summary(MLR_models[['REP']])
+```
 

### 2024-10-22 scripts/06-model_data.R

@@ -10,7 +10,10 @@
 
 #### Workspace setup ####
 library(tidyverse)
-#library(rstanarm)
+library(rstanarm)
+library(modelsummary)
+library(rstanarm)
+library(splines)
 
 #### Read data ####
 cleaned_data <- read_csv("data/02-analysis_data/cleaned_data.csv")
@@ -57,7 +60,7 @@ for (party in names(models)) {
 }
 
 # choose a party to check summary
-model_summary <- summary(models[['REP']])
+model_summary <- summary(models[['DEM']])
 # Access the R-squared and Adjusted R-squared values
 r_squared <- model_summary$r.squared
 adjusted_r_squared <- model_summary$adj.r.squared
@@ -109,7 +112,7 @@ modelsummary(models = models)
 
 # Bayesian model section 
 model_formula_dem <- cbind(num_party, sample_size - num_party) ~ log(sample_size) +
-  pollscore + numeric_grade + (1 | pollster)
+  pollscore + numeric_grade + transparency_score + days_since_start + (1 | pollster)
 
 # Set normal priors for the coefficients
 priors <- normal(0, 2.5, autoscale = TRUE)
@@ -124,6 +127,7 @@ bayesian_model_dem <- stan_glmer(
   cores = 4,
   adapt_delta = 0.95
 )
+
 priors |>
   ggplot(aes(x = marathon_time)) +
   geom_histogram(binwidth = 10) +
@@ -135,7 +139,7 @@ pp_check(bayesian_model_dem)
 plot(bayesian_model_dem, pars = "(Intercept)", prob = 0.95)
 
 spline_model_dem <- stan_glm(
-  pct ~ ns(days_since_start, df = 7) + pollscore + log(sample_size) + pollster,  # Spline with 5 degrees of freedom for end_date_num
+  pct ~ ns(days_since_start, df = 7) + pollscore + numeric_grade + transparency_score + log(sample_size) + pollster,  # Spline with 5 degrees of freedom for end_date_num
   data = dem_data,                            # DEM national dataset
   family = gaussian(),                        # Use a Gaussian family for continuous outcome
   prior = normal(0, 5),                       # Weakly informative prior for coefficients
@@ -147,7 +151,7 @@ spline_model_dem <- stan_glm(
 )
 
 spline_model_rep <- stan_glm(
-  pct ~ ns(days_since_start, df = 7) + pollscore + log(sample_size) + pollster,  # Spline with 5 degrees of freedom for end_date_num
[…]

### 2024-10-22 paper/paper.qmd
update paper

@@
 library(tidyverse)
+library(rstanarm)
+library(splines)
 ```
@@
 &+ \beta_4 \, \text{transparency\_score}_i + \beta_5 \, \text{days\_since\_start}_i + \alpha_j
-\end{align}
 \alpha_j &\sim \text{Normal}(0, \sigma_{\text{pollster}}) \\
@@
 ### Result
+Summary of Bayesian model is hsown in [@tbl-baymodelresults], and the ppcheck is in appendix [@fig-ppcheck]. 
 ```{r}
@@
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm.
-We use priors $\text{Normal}(0, 5) \\$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
+We use priors $\text{Normal}(0, 5)$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
 Then we crate a data frame like [@tbl-newdata], we set the quality scale of predict data to the best to perform a poll that has high quality.
@@
 
-posterior_preds_dem <- posterior_predict(spline_model_dem, newdata = new_data)
-posterior_preds_rep <- posterior_predict(spline_model_rep, newdata = new_data)
+posterior_preds_dem <- posterior_predict(spline_dem, newdata = new_data)
+posterior_preds_rep <- posterior_predict(spline_rep, newdata = new_data)
 
 combined_data <- bind_rows(dem_data, rep_data)
+
+# Summarize predictions
+pred_summary_dem <- new_data |>
+  mutate(
+    pred_mean_dem = colMeans(posterior_preds_dem),
+    pred_lower_dem = apply(posterior_preds_dem, 2, quantile, probs = 0.025),
+    pred_upper_dem = apply(posterior_preds_dem, 2, quantile, probs = 0.975),
+    party = "DEM"
+  )
+
+# Summarize predictions for REP
+pred_summary_rep <- new_data |>
+  mutate(
+    pred_mean_rep = colMeans(posterior_preds_rep),
+    pred_lower_rep = apply(posterior_preds_rep, 2, quantile, probs = 0.025),
+    pred_upper_rep = apply(posterior_preds_rep, 2, quantile, probs = 0.975),
+    party = "REP"
+  )
+
+# Combine both summaries into a single dataframe
+pred_summary <- bind_rows(pred_summary_dem, pred_summary_rep)
 
@@
     data = pred_summary %>% filter(party == "DEM"),
-    aes(x = days_since_start, y = pred_mean_dem, color = "DEM"),  # Move color inside aes()
+    aes(x = days_since_start, y = pred_mean_dem, color = "DEM"),
     inherit.aes = FALSE
@@
     data = pred_summary %>% filter(party == "DEM"),
-    aes(x = days_since_start, ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"),  # Move fill inside aes()
+    aes(x = days_since_start, ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"),
     alpha = 0.2,
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
 ```{r}
@@
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
+#| label: fig-ppcheck
 #| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| fig-cap: "Posterior prediction check for Bayesian model"
 
-# pp_check(first_model) +
-#   theme_classic() +
-#   theme(legend.position = "bottom")
-# 
-# posterior_vs_prior(first_model) +
-#   theme_minimal() +
-#   scale_color_brewer(palette = "Set1") +
-#   theme(legend.position = "bottom") +
-#   coord_flip()
+pp_check(Bay_model)
+
 ```

### 2024-10-22 scripts/03-clean_data.R
merge paper.qmd update cleaned data

@@ -33,7 +33,7 @@ cleaned_data <-
   raw_data |>
   janitor::clean_names() %>%
   select(pollster,sample_size,numeric_grade,pollscore,state,transparency_score,end_date,party,answer,pct) %>%
-  filter(end_date > as.Date("2024-01-01")) %>%
+  filter(end_date >= as.Date("2024-01-01")) %>%
   mutate(
     state = if_else(is.na(state), "National", state))
 

### 2024-10-22 scripts/03-clean_data.R
updated preamble

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Cleans the raw poll predicts data of United States president election 2024 from Five Thirty Eight 
+# Author: Ziqi Zhu, Yuanchen Miao, Claire Chang
+# Date: 22 October 2024
+# Contact: ziqi.zhu@mail.utoronto.ca, vincent.miao@mail.utoronto.ca, claire.chang@mail.utoronto.ca
+# License: None
+# Pre-requisites: raw data must be obtained from poll-of-polls
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-10-22 scripts/04-test_cleaned_data.R

@@ -1,14 +1,12 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-#electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
+# Purpose: Tests the structure and validity of the cleaned United States president election 2024 data
+# Author: Ziqi Zhu, Yuanchen Miao, Claire Chang
+# Date: 22 October 2024
+# Contact: ziqi.zhu@mail.utoronto.ca, vincent.miao@mail.utoronto.ca, claire.chang@mail.utoronto.ca
+# License: None
 # Pre-requisites: 
 # - The `tidyverse` package must be installed and loaded
-# - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# - 03-clean_data.R must be run
 
 
 #### Workspace setup ####

### 2024-10-22 scripts/06-model_data.R

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
+# Purpose: SLM of specific candidate are fitted and MLM of all candidates from United States president election 2024 are fitted
+# Author: Ziqi Zhu, Yuanchen Miao, Claire Chang
+# Date: 22 October 2024
+# Contact: ziqi.zhu@mail.utoronto.ca, vincent.miao@mail.utoronto.ca, claire.chang@mail.utoronto.ca
+# License: None
+# Pre-requisites: 03-clean-data.R must be run to obtained the data used in this script,04-test_cleaned_data.R must be run to verify that cleaned data could be used to fit a model
 # Any other information needed? [...UPDATE THIS...]
 
 

### 2024-10-22 paper/paper.qmd
merge paper.qmd update cleaned data
update appendix graph position

@@
 
-## SLR Model set-up
-
 ## MLR Model set-up
@@
 The $r^2$ shows that the model explain the highest amount of variance on Democratic Party than others with value 0.357, and other parties less than 0.2.
-Based on the summary of models for DEM and REP [@tbl-mlrmodelsum], we are not sure if those predictor that measure quality of polls are significant or not, so we keep them for now in Bayesian model.
+Based on the summary of models for DEM and REP [@sec-model-sigcheck], we are not sure if those predictor that measure quality of polls are significant or not, so we keep them for now in Bayesian model.
 Prediction of MLR models is shown in [@fig-mlrpredict], which we can see blue line which is Democratic party has a steeper slope and start to take the lead between day 100 and day 200.
@@
 ```{r}
+#| echo: false
 #| eval: true
-#| echo: false
-#| message: false
 #| warning: false
-#| label: fig-assumptioncheck
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: [""]
+#| fig.cap: "Residuals vs Fitted for DEM Party"
 
-for (party in names(MLR_models)) {
-  # Extract the model for the current party
-  current_model <- MLR_models[[party]]
-
-  # Create a Residuals vs Fitted plot for the current model
-  plot(current_model, which = 1, main = paste("Residuals vs Fitted for", party, "Party"))
-
-}
+plot(MLR_models[['DEM']], which = 1, main = "Residuals vs Fitted for DEM Party")
 ```
 
-## Significant check for MLR models
 ```{r}
@@
 #| eval: true
-#| label: tbl-mlrmodelsum
-#| tbl-cap: "Summary of DEM and REP data fit by MLR model"
+#| warning: false
+#| fig.cap: "Residuals vs Fitted for REP Party"
+
+plot(MLR_models[['REP']], which = 1, main = "Residuals vs Fitted for REP Party")
+```
+## Significant check for MLR models {#sec-model-sigcheck}
+see [@tbl-mlrmodelsum-dem], and [@tbl-mlrmodelsum-rep]
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-mlrmodelsum-dem
+#| tbl-cap: "Summary of DEM data fit by MLR model"
 #| warning: false
@@
 summary(MLR_models[['DEM']])
+```
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-mlrmodelsum-rep
+#| tbl-cap: "Summary of REP data fit by MLR model"
+#| warning: false
+
 summary(MLR_models[['REP']])
@@
 
-## Posterior predictive check
+\newpage
+## Posterior predictive check {#sec-model-ppcheck}
 
@@
 #| label: fig-ppcheck
-#| layout-ncol: 2
 #| fig-cap: "Posterior prediction check for Bayesian model"
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+<!-- @fig-stanareyouokay-1 is a trace plot. It shows... This suggests... -->
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+<!-- @fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests... -->
 

### 2024-10-22 paper/paper.qmd
fix typo in paper

@@
 
-The goal of our modelling strategy is try to capture the. Firstly,...
+The goal of our modelling strategy is try to capture the trend of support percentage for different parties in 2024 Presidential election as much as possible.
 In the following section we briefly describe the models we used to investigate. Background details and diagnostics are included in [Appendix -@sec-model-details].
@@
 \log \left( \frac{p_i}{1 - p_i} \right) &= \beta_0 + \beta_1 \log(\text{sample\_size}_i) + \beta_2 \, \text{pollscore}_i + \beta_3 \, \text{numeric\_grade}_i \\
-&+ \beta_4 \, \text{transparency\_score}_i + \beta_5 \, \text{days\_since\_start}_i + \alpha_j
+&+ \beta_4 \, \text{transparency\_score}_i + \beta_5 \, \text{days\_since\_start}_i + \alpha_j\\
 \alpha_j &\sim \text{Normal}(0, \sigma_{\text{pollster}}) \\

### 2024-10-23 paper/paper.qmd
remove some code from starter folder, add references

@@
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = "none") +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-# analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-# 
-# analysis_data |> 
-#   ggplot(aes(x = width, y = length)) +
-#   geom_point(alpha = 0.8) +
-#   theme_minimal() +
-#   labs(x = "Wing width (mm)",
-#        y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
 ## Predictor variables {#sec-prevar}
@@
 
-Response Variable is pct. Predictors are log(sample_size), pollscore, numeric_grade, transparency_score, and days_since_start, detail can be found in [@sec-octvar], and [@sec-prevar]
+Response Variable is pct. Predictors are log(sample_size), pollscore, numeric_grade, transparency_score, and days_since_start, detail can be find in [@sec-octvar], and [@sec-prevar]
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm.
+We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm].
 We use priors $\text{Normal}(0, 5)$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
@@
 ```
-The prediction graph contains prediction of spline fit model for two parties, blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
-We can tell that the predict support of DEM party is taking the lead by aroud 2 percent until the last day of the dataset.
+The prediction graph [@fig-splpredict] contains prediction of spline fit model for two parties, blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
+We can tell that the predicted support of DEM party is taking the lead by aroud 2 percent until the last day of the dataset.
 

### 2024-10-23 scripts/05-exploratory_data_analysis.R

@@ -15,7 +15,7 @@ library(tidyverse)
 #### Read data ####
 poll_data <- read_csv("data/02-analysis_data/cleaned_data.csv")
 candidate_avg_pct <- poll_data %>%
-  group_by(candidate_name) %>%
+  group_by(answer) %>%
   summarize(average_pct = mean(pct, na.rm = TRUE))
 
 # View the result
@@ -23,19 +23,6 @@ print(candidate_avg_pct, n=60)
 top_10_candidates <- candidate_avg_pct %>%
   arrange(desc(average_pct)) %>%  # Sort by average_pct in descending order
   head(10)
-
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
 
 #### Save model ####
 saveRDS(

### 2024-10-23 paper/paper.qmd
new render pdf

@@
 
-To get a prediction by the Vayesian model we have, we first spline fit the model, and then create a data frame and generate posterior predictions by the new data frame.
+To get a prediction by the Bayesian model we have, we first spline fit the model, and then create a data frame and generate posterior predictions by the new data frame.
+
 ## Spline Fit for Bayesian model set-up
-
 We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm].

### 2024-11-02 scripts/03-clean_data.R
change raw_data to latest

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw poll predicts data of United States president election 2024 from Five Thirty Eight 
 # Author: Ziqi Zhu, Yuanchen Miao, Claire Chang
-# Date: 22 October 2024
+# Date: 2  November 2024
 # Contact: ziqi.zhu@mail.utoronto.ca, vincent.miao@mail.utoronto.ca, claire.chang@mail.utoronto.ca
 # License: None
 # Pre-requisites: raw data must be obtained from poll-of-polls

### 2024-11-02 scripts/05-exploratory_data_analysis.R

@@ -18,11 +18,15 @@ candidate_avg_pct <- poll_data %>%
   group_by(answer) %>%
   summarize(average_pct = mean(pct, na.rm = TRUE))
 
-# View the result
+# View the candidate
 print(candidate_avg_pct, n=60)
 top_10_candidates <- candidate_avg_pct %>%
   arrange(desc(average_pct)) %>%  # Sort by average_pct in descending order
   head(10)
+
+# View parties
+nationwide_data <- poll_data %>%  filter(state == "National")
+boxplot(pct ~ party, data = nationwide_data)
 
 #### Save model ####
 saveRDS(

### 2024-11-02 scripts/06-model_data.R

@@ -128,18 +128,13 @@ bayesian_model_dem <- stan_glmer(
   adapt_delta = 0.95
 )
 
-priors |>
-  ggplot(aes(x = marathon_time)) +
-  geom_histogram(binwidth = 10) +
-  theme_classic()
-
 # Posterior predictive checks
 pp_check(bayesian_model_dem)
 
 plot(bayesian_model_dem, pars = "(Intercept)", prob = 0.95)
 
 spline_model_dem <- stan_glm(
-  pct ~ ns(days_since_start, df = 7) + pollscore + numeric_grade + transparency_score + log(sample_size) + pollster,  # Spline with 5 degrees of freedom for end_date_num
+  pct ~ ns(days_since_start, df = 7) + pollscore + numeric_grade + transparency_score + log(sample_size) + pollster,  # Spline with 7 degrees of freedom for end_date_num
   data = dem_data,                            # DEM national dataset
   family = gaussian(),                        # Use a Gaussian family for continuous outcome
   prior = normal(0, 5),                       # Weakly informative prior for coefficients
@@ -151,94 +146,20 @@ spline_model_dem <- stan_glm(
 )
 
 spline_model_rep <- stan_glm(
-  pct ~ ns(days_since_start, df = 7) + pollscore + numeric_grade + transparency_score + log(sample_size) + pollster,  # Spline with 5 degrees of freedom for end_date_num
-  data = rep_data,                            # DEM national dataset
-  family = gaussian(),                        # Use a Gaussian family for continuous outcome
-  prior = normal(0, 5),                       # Weakly informative prior for coefficients
-  prior_intercept = normal(50, 10),           # Weakly informative prior for intercept
-  seed = 1234,                                # Set seed for reproducibility
-  iter = 2000,                                # Number of iterations
-  chains = 4,                                 # Number of MCMC chains
-  refresh = 0                                 # Suppress unnecessary output
+  pct ~ ns(days_since_start, df = 7) + pollscore + numeric_grade + transparency_score + log(sample_size) + pollster,
+  data = rep_data,
+  family = gaussian(),
+  prior = normal(0, 5),
+  prior_intercept = normal(50, 10),
+  seed = 1234,
[…]

### 2024-11-03 paper/paper.qmd
Updated idealized methodology and added a survey

@@
 The goal of our modelling strategy is try to capture the trend of support percentage for different parties in 2024 Presidential election as much as possible.
-In the following section we briefly describe the models we used to investigate. Background details and diagnostics are included in [Appendix -@sec-model-details].
+In the following section we briefly describe the models we used to investigate.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
 ## MLR Model set-up
+
 We first filtered our data by different parties, as the single linear model but include other 3 parties(GRE, IND, and LIB, Conservative only has two entries in cleaned nationwide data and has 0 pct so its been removed), and fit linear model for each of them.
-The formula for Multiple linear regression is as follow
-\begin{align}
+The formula for Multiple linear regression is as follow \begin{align}
 \text{pct}_i &= \beta_0 + \beta_1 \log(\text{sample\_size}_i) + \beta_2 \, \text{pollscore}_i + \beta_3 \, \text{numeric\_grade}_i \\
@@
 ```
+
 We first check the residual vs.
@@
 ## Bayesian Model set-up
+
 The Bayesian model is build from the MLR models from above section.
@@
 
+```{=tex}
 \begin{align} 
@@
 \end{align}
-
+```
 $y_i$ represent the number of individuals in the sample that support the Democratic party (this corresponds to num_party).
@@
 ### Result
+
 Summary of Bayesian model is hsown in [@tbl-baymodelresults], and the ppcheck is in appendix [@fig-ppcheck].
+
 ```{r}
@@
 ## Spline Fit for Bayesian model set-up
+
 We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm].
@@
 the result is shown as [@sec-splresult]
+
 ```{r}
@@
 ## Result {#sec-splresult}
+
 ```{r}
@@
 ```
+
 The prediction graph [@fig-splpredict] contains prediction of spline fit model for two parties, blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 
-## Idealized  Survey and Methodology
+## Idealized Methodology
+
+To forecast the 2024 U.S.
+presidential election with a budget of $100,000, this methodology combines stratified sampling, multimodal recruitment, and aggregation techniques that leverage the strengths of multiple data sources.
+The approach employs stratified random sampling with targeted oversampling of key subgroups, like younger and minority voters, to address their frequent underrepresentation in polling.
+By segmenting eligible U.S.
+voters by demographics—age, gender, race, party affiliation, region, and urban/rural residency—the method ensures broad representational coverage, aiming to capture diverse voting patterns accurately [@pewpolling] .
+Partnering with a voter database provider gives us access to a detailed list of registered voters, organized by demographics, to improve the accuracy of our sample.
+This approach also builds on public trust, as Americans generally trust polls from news organizations (43%) more than those from websites that combine multiple polls (30%) [@PublicOpinionQuarterly].
+
+Respondents are recruited through multiple channels to reduce potential biases associated with any single mode.
+The multimodal approach includes Interactive Voice Response (IVR) to reach older and rural populations (30%), targeted online surveys to capture difficult-to-reach demographics (40%), text surveys directed at younger voters (20%), and live calls in key battleground states (10%).
+This mix helps address “coverage error” by compensating for the limitations associated with each mode [@Blumenthal_2014].
+
+Survey data is collected via a short, 5–7 minute questionnaire with 10 core questions covering demographics, voting intention, likelihood of voting, and key issues.
+A shorter survey length and mix of direct and indirect questions help reduce Social Desirability Bias, while randomizing question order minimizes positional bias.
+Given that any individual survey is likely to suffer from random and systematic errors—sampling errors, coverage errors, and response biases—aggregation across multiple polls can mitigate these sources of error, improving the accuracy of forecasts by averaging out errors specific to individual polls [@Blumenthal_2014].
+
+For data validation and quality control, screening questions confirm attentiveness, demographic cross-checks validate responses against voter registration data, and post-survey weighting adjusts for demographic imbalances.
+Weighting responses to match population demographics ensures that key groups are proportionally represented, while data cleaning removes low-quality responses to maintain data integrity [@Blumenthal_2014].
+To further address biases, surveys are conducted anonymously in formats like online and SMS.
+Weighting adjustments are also made to address non-response bias, enhancing accuracy by reflecting the true population distribution.
+We will keep the process and result transparent by posting on our website for the public to view.
+
+Weekly aggregation serves as a “poll of polls” that balances fluctuations across samples and smooths out errors.
+By incorporating recent data more heavily, this rolling average adapts to shifts in voter sentiment leading up to the election.
+Bayesian modeling, applied to smooth sample variations and incorporate prior election trends, stabilizes the forecast further.
+Poll aggregation, which is commonly employed in election analyses, effectively combines estimates from various samples, reducing random error and discounting non-universal biases [@Blumenthal_2014].
+Given the variability in survey methodologies—such as IVR surveys missing cell-only populations or web surveys excluding offline individuals—aggregation allows errors in one type of survey to offset those in another.
+Weighting each survey according to sample size and precision further refines the accuracy of the aggregated forecast.
+
+### Budget Allocation:
+
+- **Sample Acquisition**: Partnership with a voter database provider to access a list of registered voters across all states $10,000
+
+- **Interactive Voice Response (IVR)**: Automated voice surveys for older and rural demographics $20,000
+
+- **Online Surveys**: Targeted ads and opt-in digital surveys to capture harder-to-reach groups (40% of sample) $25,000
+
+- **Text Messaging**: SMS-based surveys to engage younger demographics (20% of sample) $10,000
+
+- **Live Phone Calls**: Calls in battleground states to reach underrepresented groups (10% of sample) $20,000
+
+- **Data Validation & Cleaning**:	Screening questions, demographic verification, and removal of low-quality responses to maintain data integrity $5,000
+
+- **Data Analysis & Forecast Modeling**	Bayesian modeling, poll aggregation, and analysis of survey trends to produce election forecasts $10,000
+
+## Idealized Survey
+
+The proposed survey questionnarie design is in the following link: https://forms.gle/Zm5Kfj3kL58gwCz38
+
+### Survey Copy
+
+1. **What is your age??**
+   - 18-29
+   - 30-44
+   - 45-64
+   - 65+
+
+2. **What is your gender identity?**
+   - Female
+   - Male
+   - Another Gender Identity
+
+3. **Which of the following best describes your race/ethnicity?**
+   - White
+   - Black/African American
+   - Hispanic/Latino
+   - Asian
+   - Native American
+   - Prefer not to say
+   - Other
+
+4. **What is your highest level of education?**
+   - High school or less
+   - College
+   - Bachelor's degree
+   - Graduate degree
+   - Prefer not to say
+
+5. **In which U.S. region do you currently reside?**
+   - Midwest
+   - Northeast
+   - South
+   - West
+
+6. **Are you registered to vote in the 2024 presidential election?**
+   - Yes
+   - No
+
+7. **2024 Which candidate do you plan to vote for?**
+   - Kamala Harris
+   - Donald Trump
+   - Undecided
+   - Other
+
+8. **2020 Which candidate did you vote for?**
+   - Joe Biden
+   - Donald Trump
+   - Did not vote
+   - Prefer not to say
+   - Other
+
+9. **What is the most important issue influencing your vote?**
+   - Economy
+   - Healthcare 
+   - Immigration
+   - Climate Change
+   - Social Justice
+   - National Security
+   - Other
+
+10. **On a scale of 1-5, how likely are you to vote in the 2024 U.S. presidential election?**
+
+   - 1 being Not Likely to Vote
+   - 5 being Likely to Vote
 
@@
 The population consists of all eligible voters in the U.S., while the sampling frame includes registered voters across states, segmented by demographics like age, gender, party affiliation, and ethnicity.
-Trafalgar Group typically samples likely voters, focusing on those most likely to participate in upcoming elections.
-Trafalgar Group recruits its sample using a mix of interactive voice response (IVR), live phone calls, text messages, online panels, and email surveys.
-Trafalgar Group employs stratified random sampling to ensure proportional representation of subgroups like political party, gender, and region.
-This method improves accuracy by reflecting the electorate's demographic and political makeup but may risk over-stratification, giving smaller voter groups disproportionate influence and potentially skewing results.
-Trafalgar Group handles non-response by using weighting to adjust the sample, ensuring respondent demographics match population parameters.
-This method reduces bias by accounting for underrepresented groups, though heavy reliance on post-survey adjustments may introduce new biases, particularly when there are large discrepancies in response rates across demographics.
+Trafalgar Group typically samples likely voters,focusing on those most likely to participate in upcoming elections.
+Trafalgar Group recruits its sample using a mix of interactive voice response, live phone calls, text messages, emails, digital dial back interface and online targeted opt-in digital survey platforms [@trafalgar_polling_methodology].
+Trafalgar places particular emphasis on reducing Social Desirability Bias, designing short questionnaires and using nontraditional question formats to help participants feel more comfortable expressing their true preferences, especially on sensitive topics [@trafalgar_polling_methodology].
+They also adjust for non-response by weighting the sample, ensuring it reflects the broader population’s demographic composition.
+While this helps to reduce bias from underrepresented groups, reliance on post-survey adjustments can introduce new biases, especially if response rates vary significantly among demographic groups.
 
 # Model details {#sec-model-details}
+
 ## Assumption check for MLR models
+
 ```{r}
@@
 ```
+
 ## Significant check for MLR models {#sec-model-sigcheck}
+
 see [@tbl-mlrmodelsum-dem], and [@tbl-mlrmodelsum-rep]
+
 ```{r}
@@
 ```
+
 ```{r}
@@
 \newpage
+
 ## Posterior predictive check {#sec-model-ppcheck}

### 2024-11-03 paper/paper.qmd
Intro is updated

@@
 This model provides a practical tool for researchers, political strategists, and analysts to assess the reliability of polling data and its implications for elections.
+
+This paper develops statistical models to forecast the 2024 presidential election outcome between Kamala Harris and Donald Trump.
+Utilizing multiple linear regression, we predict the support percentages for each candidate across various states, incorporating essential factors such as pollster rating, sample size, and state-specific demographics.
+By aggregating these state-level predictions, we simulate possible outcomes to estimate each candidate's probability of achieving the necessary 270 electoral votes.
+Additionally, we analyze [@trafalgar_polling_methodolog] polling methodology and propose an optimized survey and methodology aimed at improving election forecast accuracy.
+
+The remainder of this paper is structured as follows.
+@sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
+@sec-model outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
+@sec-predict presents our Electoral College predictions based on the model outputs.
+@sec-discuss discusses the implications of our findings and suggests directions for future research.
+Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.
 
@@
 By segmenting eligible U.S.
-voters by demographics—age, gender, race, party affiliation, region, and urban/rural residency—the method ensures broad representational coverage, aiming to capture diverse voting patterns accurately [@pewpolling] .
+voters by demographics—age, gender, race, party affiliation, region, and urban/rural residency—the method ensures broad representational coverage, aiming to capture diverse voting patterns accurately [@pewpolling].
 Partnering with a voter database provider gives us access to a detailed list of registered voters, organized by demographics, to improve the accuracy of our sample.

### 2024-11-03 paper/paper.qmd
Intro updated

@@
 By aggregating these state-level predictions, we simulate possible outcomes to estimate each candidate's probability of achieving the necessary 270 electoral votes.
-Additionally, we analyze [@trafalgar_polling_methodolog] polling methodology and propose an optimized survey and methodology aimed at improving election forecast accuracy.
+Additionally, we analyze [@trafalgar_polling_methodology] polling methodology and propose an optimized survey and methodology aimed at improving election forecast accuracy.
 
@@
 @sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
-@sec-model outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
-@sec-predict presents our Electoral College predictions based on the model outputs.
+@sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
+@sec-predict presents our *** predictions based on the model outputs.
 @sec-discuss discusses the implications of our findings and suggests directions for future research.
@@
 
-Our data is download from [Five Thirty Eight](https://projects.fivethirtyeight.com/polls)[@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate**[@lubridate], **dplyr**[@dplyr], **tidyverse**[@tidyverse], **model summary**[@models] to process the data.
+Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models] to process the data.
 Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
-Graphs are created by using ggplot2 in from package **tidyverse**[@tidyverse]Following **Telling stories with data**[@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
+Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse]Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
 We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
@@
 
-# Models
+# Models {#sec-models}
 
@@
 
-# Prediction
+# Prediction {#sec-predict}
 
@@
 
-# Discussion
+# Discussion {#sec-discuss}
 

### 2024-11-03 paper/paper.qmd
add bayesian model for REP and predict results

@@
 The response is modeled as binomial: $y_i \sim \text{Binomial}(\text{sample\_size}_i, p_i)$, where $p_i$ is the probability that an individual in poll $i$ supports the Democratic party, and $\text{sample\_size}_i$ is the total number of individuals surveyed in poll $i$.
+We also create a model for Republican to compare the trend in prediction result in the following section [@sec-bayresult].
 
-### Result
-
-Summary of Bayesian model is hsown in [@tbl-baymodelresults], and the ppcheck is in appendix [@fig-ppcheck].
+### Result {#sec-bayresult}
+To get a prediction by the Bayesian model, we create a data frame and generate posterior predictions by the new data frame, example data looks like [@tbl-newdata], we set the quality scale of predict data to the best to perform a poll that has high quality.
+Summary of Bayesian model is shown in [@tbl-baymodelresults], and the posterior predictive check is in appendix [@sec-model-ppcheck].
+The predict result is shown in [@fig-baypredict].
 
@@
 
-Bay_model <-
+bayesian_model_dem <-
   readRDS(file = here::here("models/bayesian_model_dem.rds"))
+bayesian_model_rep <-
+  readRDS(file = here::here("models/bayesian_model_rep.rds"))
 ```
@@
 modelsummary::modelsummary(
-  Bay_model,
+  bayesian_model_dem,
   fmt = 2
@@
 ```
-
-# Prediction {#sec-predict}
-
-To get a prediction by the Bayesian model we have, we first spline fit the model, and then create a data frame and generate posterior predictions by the new data frame.
-
-## Spline Fit for Bayesian model set-up
-
-We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm].
-We use priors $\text{Normal}(0, 5)$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
-Then we crate a data frame like [@tbl-newdata], we set the quality scale of predict data to the best to perform a poll that has high quality.
-the result is shown as [@sec-splresult]
 
@@
 new_data <- data.frame(
+  num_party = 0,
   pollscore = -1.5,
@@
   ),
-  pollster = factor("TIPP", levels = levels(dem_data$pollster))
+  pollster = factor("YouGov", levels = levels(dem_data$pollster))
 )
@@
 ```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+set.seed(123) # set seed for reproduce of predict result
+predict_counts_dem <- posterior_predict(bayesian_model_dem, newdata = new_data)
+predict_counts_rep <- posterior_predict(bayesian_model_rep, newdata = new_data)
+
+combined_data <- bind_rows(dem_data, rep_data)
+pred_summary_dem <- new_data %>%
+  mutate(
+    pred_mean_dem = colMeans(predict_counts_dem),
+    pred_lower_dem = apply(predict_counts_dem, 2, quantile, probs = 0.025),
+    pred_upper_dem = apply(predict_counts_dem, 2, quantile, probs = 0.975)
+  )
+
+# Summarize predictions for REP
+pred_summary_rep <- new_data %>%
+  mutate(
+    pred_mean_rep = colMeans(predict_counts_rep),
+    pred_lower_rep = apply(predict_counts_rep, 2, quantile, probs = 0.025),
+    pred_upper_rep = apply(predict_counts_rep, 2, quantile, probs = 0.975)
+  )
+
+# Combine both summaries
+combined_summary <- pred_summary_dem %>%
+  select(days_since_start, pred_mean_dem, pred_lower_dem, pred_upper_dem) %>%
+  left_join(
+    pred_summary_rep %>% 
+      select(days_since_start, pred_mean_rep, pred_lower_rep, pred_upper_rep),
+    by = "days_since_start"
+  )
+# Extract the last day data for annotations
+last_day_data <- combined_summary %>%
+  filter(days_since_start == max(days_since_start))
+```
+
+```{r}
+#| echo: false
+#| label: fig-baypredict
+#| fig-cap: "Predicted number of supporters for DEM and REP party"
+#| warning: false
+
+ggplot(combined_summary, aes(x = days_since_start)) +
+  geom_line(aes(y = pred_mean_dem, color = "DEM"), size = 1) +
+  geom_ribbon(aes(ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"), alpha = 0.2) +
+  geom_line(aes(y = pred_mean_rep, color = "REP"), size = 1) +
+  geom_ribbon(aes(ymin = pred_lower_rep, ymax = pred_upper_rep, fill = "REP"), alpha = 0.2) +
+  scale_color_manual(values = c("DEM" = "blue", "REP" = "red")) +
+  scale_fill_manual(values = c("DEM" = "blue", "REP" = "red")) +
+  geom_text(
+    data = last_day_data,
+    aes(
+      x = days_since_start,
+      y = pred_mean_dem,
+      label = paste("DEM:", round(pred_mean_dem))
+    ),
+    color = "blue",
+    vjust = -1,
+    hjust = 1
+  ) +
+  geom_text(
+    data = last_day_data,
+    aes(
+      x = days_since_start,
+      y = pred_mean_rep,
+      label = paste("REP:", round(pred_mean_rep))
+    ),
+    color = "red",
+    vjust = -1,
+    hjust = 1
+  ) +
+  labs(
+    x = "Days Since Start",
+    y = "Predicted Number of Supporters",
+    color = "Party",
+    fill = "Party"
+  ) +
+  theme_minimal()
+```
+
+The prediction results of the Bayesian model closely align with those of the linear model, with both showing the mean predicted number of supporters for each party.
+Notably, the Democratic Party takes the lead around day 210 (late July).
+By early November, the predictions indicate that the Democratic Party has an average of 601 supporters, compared to the Republican Party's 578.
+This suggests a slight advantage for the Democrats as the final prediction period concludes.
+
+## Spline Fit model set-up
+
+We first run the model in R [@citeR] using the `rstanarm` package of [@rstanarm], and use priors $\text{Normal}(0, 5)$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
+Then we use the same data frame that used for predict for Bayesian model [@tbl-newdata], the result is shown as [@sec-splresult]
 
@@
 #| echo: false
-#| label: fig-splpredict
-#| fig-cap: "Poll Percentage over Time with Spline Fit for DEM and REP party"
 #| warning: false
+#| message: false
 
@@
 
-# Plot the spline fit
-ggplot(combined_data, aes(x = days_since_start, y = pct, color = pollster)) +
-  # DEM predictions
-  geom_line(
-    data = pred_summary %>% filter(party == "DEM"),
-    aes(x = days_since_start, y = pred_mean_dem, color = "DEM"),
-    inherit.aes = FALSE
+# Extract the last day data for annotations
+last_day_data <- pred_summary %>%
+  filter(days_since_start == max(days_since_start))
+```
+
+```{r}
+#| echo: false
+#| label: fig-splpredict
+#| fig-cap: "Predicted support percentage over Time with Spline Fit for DEM and REP party"
+#| warning: false
+
+ggplot(pred_summary, aes(x = days_since_start)) +
+  geom_line(aes(y = pred_mean_dem, color = "DEM"), size = 1, data = subset(pred_summary, party == "DEM")) +
+  geom_ribbon(aes(ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"), alpha = 0.2, data = subset(pred_summary, party == "DEM")) +
+  geom_line(aes(y = pred_mean_rep, color = "REP"), size = 1, data = subset(pred_summary, party == "REP")) +
+  geom_ribbon(aes(ymin = pred_lower_rep, ymax = pred_upper_rep, fill = "REP"), alpha = 0.2, data = subset(pred_summary, party == "REP")) +
+  scale_color_manual(values = c("DEM" = "blue", "REP" = "red")) +
+  scale_fill_manual(values = c("DEM" = "blue", "REP" = "red")) +
+    geom_text(
+    data = last_day_data,
+    aes(
+      x = days_since_start,
+      y = pred_mean_dem,
+      label = paste("DEM:", round(pred_mean_dem, 2))
+    ),
+    color = "blue",
+    vjust = -1,
+    hjust = 1
   ) +
-  geom_ribbon(
-    data = pred_summary %>% filter(party == "DEM"),
-    aes(x = days_since_start, ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"),
-    alpha = 0.2,
-    inherit.aes = FALSE
+  geom_text(
+    data = last_day_data,
+    aes(
+      x = days_since_start,
+      y = pred_mean_rep,
+      label = paste("REP:", round(pred_mean_rep, 2))
+    ),
+    color = "red",
+    vjust = -1,
+    hjust = 1
   ) +
-  # REP predictions
-  geom_line(
-    data = pred_summary %>% filter(party == "REP"),
-    aes(x = days_since_start, y = pred_mean_rep, color = "REP"),  # Move color inside aes()
-    inherit.aes = FALSE
-  ) +
-  geom_ribbon(
-    data = pred_summary %>% filter(party == "REP"),
-    aes(x = days_since_start, ymin = pred_lower_rep, ymax = pred_upper_rep, fill = "REP"),  # Move fill inside aes()
-    alpha = 0.2,
-    inherit.aes = FALSE
-  ) +
-  # Use scale_color_manual and scale_fill_manual for colors and fills
-  scale_color_manual(values = c("DEM" = "blue", "REP" = "red"), 
-                     name = "Party",
-                     labels = c("DEM (Blue)", "REP (Red)")) +
-  scale_fill_manual(values = c("DEM" = "blue", "REP" = "red"), 
-                    name = "Party",
-                    labels = c("DEM (Blue)", "REP (Red)")) +
   labs(
-    x = "Days since 2024",
-    y = "Percentage",
+    x = "Days Since Start",
+    y = "Predicted Number of Supporters",
+    color = "Party",
+    fill = "Party"
   ) +
@@
 The prediction graph [@fig-splpredict] contains prediction of spline fit model for two parties, blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
-We can tell that the predicted support of DEM party is taking the lead by aroud 2 percent until the last day of the dataset.
+We can tell that the predicted support of DEM party is taking the lead by around 2 percent until the last day of the dataset.
 
@@
 #| label: fig-ppcheck
-#| fig-cap: "Posterior prediction check for Bayesian model"
+#| fig-cap: "Posterior prediction check for Bayesian models"
+#| layout-ncol: 2
+#| fig-subcap: ["Democratic Model", "Republican Model"]
 
-pp_check(Bay_model)
-
+pp_check(bayesian_model_dem)
+pp_check(bayesian_model_rep)
 ```

### 2024-11-03 scripts/05-exploratory_data_analysis.R

@@ -28,10 +28,11 @@ top_10_candidates <- candidate_avg_pct %>%
 nationwide_data <- poll_data %>%  filter(state == "National")
 boxplot(pct ~ party, data = nationwide_data)
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
+# Separate by state 
+poll_counts <- poll_data %>%
+  group_by(state) %>%  # Group by 'state'
+  summarize(poll_count = n()) %>%  # Count the number of polls per state
+  arrange(desc(poll_count))
+
 
 

### 2024-11-03 scripts/06-model_data.R

@@ -111,8 +111,8 @@ ggplot(combined_data, aes(x = days_since_start)) +
 modelsummary(models = models)
 
 # Bayesian model section 
-model_formula_dem <- cbind(num_party, sample_size - num_party) ~ log(sample_size) +
-  pollscore + numeric_grade + transparency_score + days_since_start + (1 | pollster)
+model_formula_dem <- cbind(num_party, sample_size - num_party) ~ log(sample_size) + days_since_start + 
+   pollscore + numeric_grade + transparency_score + (1 | pollster)
 
 # Set normal priors for the coefficients
 priors <- normal(0, 2.5, autoscale = TRUE)
@@ -125,6 +125,16 @@ bayesian_model_dem <- stan_glmer(
   prior_intercept = priors,
   seed = 123,
   cores = 4,
+  adapt_delta = 0.95,
+)
+
+bayesian_model_rep <- stan_glmer(
+  formula = model_formula_dem,
+  data = rep_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
   adapt_delta = 0.95
 )
 
@@ -181,14 +191,15 @@ ggplot(just_harris_high_quality, aes(x = end_date)) +
   scale_y_continuous(labels = scales::percent_format(scale = 1))+
   labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
 
-
-
 #### Save model ####
 saveRDS(
   bayesian_model_dem,
   file = "models/bayesian_model_dem.rds"
 )
-
+saveRDS(
+  bayesian_model_rep,
+  file = "models/bayesian_model_rep.rds"
+)
 saveRDS(models, file = "models/MLRmodels.rds")
 saveRDS(spline_model_dem, file = "models/spline_dem.rds")
 saveRDS(spline_model_rep, file = "models/spline_rep.rds")

### 2024-11-04 paper/paper.qmd
fix some grammars

@@
 
-We first check the residual vs.
-fitted plot as shown in [Appendix @sec-model-details], and see no obvious violation of assumptions.
+First, we checked the residual vs.
+fitted plot as shown in [Appendix @sec-model-details] and saw no obvious violation of assumptions.
 According to the summary, DEM and REP models are based on the largest number of observations (both 1235).
 Smaller parties have fewer observations, which may lead to less robust models.
-The $r^2$ shows that the model explain the highest amount of variance on Democratic Party than others with value 0.357, and other parties less than 0.2.
-Based on the summary of models for DEM and REP [@sec-model-sigcheck], we are not sure if those predictor that measure quality of polls are significant or not, so we keep them for now in Bayesian model.
-Prediction of MLR models is shown in [@fig-mlrpredict], which we can see blue line which is Democratic party has a steeper slope and start to take the lead between day 100 and day 200.
+The $r^2$ shows that the model explains the highest amount of variance on the Democratic Party than others with a value of 0.357, and other parties less than 0.2.
+Based on the summary of models for DEM and REP [@sec-model-sigcheck], we are not sure if those predictors that measure the quality of polls are significant, so we keep them for now in the Bayesian model.
+The prediction of MLR models is shown in [@fig-mlrpredict], in which we can see the blue line represents the Democratic party has a steeper slope and starts to take the lead between day 100 and day 200.
 
@@
 
-The Bayesian model is build from the MLR models from above section.
-With the inspiration from the example R code, we introduce the random effects for Pollsters, and we use a logistic link to predict the probability of support, modeling count data which is in binomial distribution.
-We choose to use a default prior and enable autoscale, the formula is shown below
+The Bayesian model is built from the MLR models from the above section.
+With inspiration from the example R code, we introduce the random effects for Pollsters, and we use a logistic link to predict the probability of support, modeling count data which is in a binomial distribution.
+We choose to use a default prior and enable autoscale, the formula is shown below.
 
@@
 ```
-$y_i$ represent the number of individuals in the sample that support the Democratic party (this corresponds to num_party).
+$y_i$ represents the number of individuals in the sample that support the Democratic party (this corresponds to num_party).
 The response is modeled as binomial: $y_i \sim \text{Binomial}(\text{sample\_size}_i, p_i)$, where $p_i$ is the probability that an individual in poll $i$ supports the Democratic party, and $\text{sample\_size}_i$ is the total number of individuals surveyed in poll $i$.
-We also create a model for Republican to compare the trend in prediction result in the following section [@sec-bayresult].
+We also create a model for the Republican party to compare the trend in prediction results in the following section [@sec-bayresult].
 
 ### Result {#sec-bayresult}
-To get a prediction by the Bayesian model, we create a data frame and generate posterior predictions by the new data frame, example data looks like [@tbl-newdata], we set the quality scale of predict data to the best to perform a poll that has high quality.
-Summary of Bayesian model is shown in [@tbl-baymodelresults], and the posterior predictive check is in appendix [@sec-model-ppcheck].
-The predict result is shown in [@fig-baypredict].
+To get a prediction by the Bayesian model, we create a data frame and generate posterior predictions by the new data frame, this table [@tbl-newdata] shows the first several rows, and we set the quality scale of predicted data to the best to perform a poll that has high quality.
+A summary of the Bayesian model is shown in [@tbl-baymodelresults], and the posterior predictive check is in appendix [@sec-model-ppcheck].
+The predicted result is shown in [@fig-baypredict].
 
@@
 
-head(new_data)
+head(new_data, 3)
 ```
@@
 We first run the model in R [@citeR] using the `rstanarm` package of [@rstanarm], and use priors $\text{Normal}(0, 5)$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
-Then we use the same data frame that used for predict for Bayesian model [@tbl-newdata], the result is shown as [@sec-splresult]
+Then we use the same data frame that was used to predict for the Bayesian model [@tbl-newdata], the result is shown as [@sec-splresult]
 
@@
 
-The prediction graph [@fig-splpredict] contains prediction of spline fit model for two parties, blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
-We can tell that the predicted support of DEM party is taking the lead by around 2 percent until the last day of the dataset.
+The prediction graph [@fig-splpredict] contains a prediction of the spline fit model for two parties, the blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
+We can tell that the predicted support of the DEM party is taking the lead by around 2 percent until the last day of the dataset.
 
@@
 
-Weaknesses and next steps should also be included.
+Firstly we focused only on nationwide polls, excluding state-specific data, due to limited and outdated information from certain states.
+For example, Mississippi had only four polls, with the most recent conducted in April 2024, making state-level modeling and vote counting challenging.
+This limitation restricts the accuracy of state-by-state analysis.
+With more reliable and updated data, using a state-based approach would enhance the model’s precision and provide a more comprehensive prediction of the election outcome.
 

### 2024-11-04 paper/paper.qmd
add weakness

@@
 The increase in support for the DEM party in the predictions could be attributed to Harris's rise as the Democratic candidate after Biden's exit from the race On July 21, 2024, which is about 230 days from start of 2024.
+By analyzing the support percentages of Biden and Harris from the Democratic poll data @fig-demonly, it becomes evident that their combined impact likely contributes to the double peak seen in the posterior predictive check @fig-ppcheck.
+This dual pattern challenges the model’s ability to capture trends accurately
+
+```{r}
+#| echo: false
+#| label: fig-demonly
+#| fig-cap: "Density Plot of Support Percentage for Biden and Harris in 2024"
+#| warning: false
+
+dem_data_filtered <- dem_data %>%
+  filter(answer %in% c("Biden", "Harris"))
+
+# Create the density plot
+ggplot(dem_data_filtered, aes(x = pct, fill = answer)) +
+  geom_density(alpha = 0.4) +
+  labs(
+    x = "Support Percentage",
+    y = "Density",
+    fill = "Candidate"
+  ) +
+  theme_minimal()
+```
 
@@
 With more reliable and updated data, using a state-based approach would enhance the model’s precision and provide a more comprehensive prediction of the election outcome.
+Furthermore modeling Biden and Harris together under one Democratic model introduces potential inaccuracies, especially influencing the slope of predictions in both linear and Bayesian models, making the observed upward trend less reliable.
 
@@
 
+In @fig-ppcheck we implement a posterior predictive check.
+The check for DEM model shows the observed data has two peaks, indicating a possible bimodal distribution, but the predictive simulations also show significant variability around the peaks.
+While the model for REP performs better, capturing the main distribution of the data
+
 ```{r}
@@
 
-pp_check(bayesian_model_dem)
-pp_check(bayesian_model_rep)
+pp_check(bayesian_model_dem) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+pp_check(bayesian_model_rep) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
 ```

### 2024-11-04 scripts/05-exploratory_data_analysis.R

@@ -34,5 +34,13 @@ poll_counts <- poll_data %>%
   summarize(poll_count = n()) %>%  # Count the number of polls per state
   arrange(desc(poll_count))
 
+# Check support pct for Biden and Harris
+average_biden <- nationwide_data %>%
+  filter(answer == "Biden") %>%
+  summarize(avg_pct_biden = mean(pct, na.rm = TRUE))
 
+# Calculate average support for Harris
+average_harris <- nationwide_data %>%
+  filter(answer == "Harris") %>%
+  summarize(avg_pct_harris = mean(pct, na.rm = TRUE))
 

### 2024-11-04 paper/paper.qmd
title and abstract updated

--- 2024-11-03 paper/paper.qmd
@@
 ---
-title: "Forecasting the US Presidential Election: A Poll-of-Polls Approach Using Linear Models By Using Data of Pollsters From Five Thirty Eight In 2024"
+title: "Forecasting the 2024 U.S. Presidential Election Results: Linear Regression, Bayesian and Spline-Fit Gaussian Model by Using Pollsters' Data From Five Thirty Eight"
+subtitle: "Contrasting Harris's Lead: Strong in Linear Regression and Bayesian Model, Modest in Spline-Fit Gaussian Model"
 author: 
@@
 date-format: long
-abstract: "This paper presents a multiple linear regression model for forecasting the outcome of the 2024 U.S.
-Presidential Election.
-Using nationwide polling data, we predict the percentage of support for U.S.
-presidential candidates based on key factors such as sample size, pollster quality, and timing.
-Our model aggregates these predictions to simulate potential election outcomes.
-The results reveal how specific poll attributes affect the accuracy of forecasts, offering valuable insights for predicting the likelihood of various electoral scenarios."
+abstract: "we use FiveThirtyEight’s 2024 U.S.
+presidential poll-of-polls data to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
+By focusing on key reliability metrics—such as sample size, numeric scores, poll scores, and state-level factors—we fit the data into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
+Our findings reveal a strong lead for Harris in both the linear regression and Bayesian models, with a more modest lead in the spline-fit Gaussian model.
+These results offer insight into the potential trajectory of the election, helping the public anticipate the policy directions that might follow a Harris administration."
 number-sections: true
@@
 The response is modeled as binomial: $y_i \sim \text{Binomial}(\text{sample\_size}_i, p_i)$, where $p_i$ is the probability that an individual in poll $i$ supports the Democratic party, and $\text{sample\_size}_i$ is the total number of individuals surveyed in poll $i$.
-We also create a model for Republican to compare the trend in prediction result in the following section [@sec-bayresult].
 
-### Result {#sec-bayresult}
-To get a prediction by the Bayesian model, we create a data frame and generate posterior predictions by the new data frame, example data looks like [@tbl-newdata], we set the quality scale of predict data to the best to perform a poll that has high quality.
-Summary of Bayesian model is shown in [@tbl-baymodelresults], and the posterior predictive check is in appendix [@sec-model-ppcheck].
-The predict result is shown in [@fig-baypredict].
+### Result
+
+Summary of Bayesian model is hsown in [@tbl-baymodelresults], and the ppcheck is in appendix [@fig-ppcheck].
 
@@
 
-bayesian_model_dem <-
+Bay_model <-
   readRDS(file = here::here("models/bayesian_model_dem.rds"))
-bayesian_model_rep <-
-  readRDS(file = here::here("models/bayesian_model_rep.rds"))
 ```
@@
 modelsummary::modelsummary(
-  bayesian_model_dem,
+  Bay_model,
   fmt = 2
@@
 ```
+
+# Prediction {#sec-predict}
+
+To get a prediction by the Bayesian model we have, we first spline fit the model, and then create a data frame and generate posterior predictions by the new data frame.
+
+## Spline Fit for Bayesian model set-up
+
+We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm].
+We use priors $\text{Normal}(0, 5)$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
+Then we crate a data frame like [@tbl-newdata], we set the quality scale of predict data to the best to perform a poll that has high quality.
+the result is shown as [@sec-splresult]
 
@@
 new_data <- data.frame(
-  num_party = 0,
   pollscore = -1.5,
@@
   ),
-  pollster = factor("YouGov", levels = levels(dem_data$pollster))
+  pollster = factor("TIPP", levels = levels(dem_data$pollster))
 )
@@
 ```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-set.seed(123) # set seed for reproduce of predict result
-predict_counts_dem <- posterior_predict(bayesian_model_dem, newdata = new_data)
-predict_counts_rep <- posterior_predict(bayesian_model_rep, newdata = new_data)
-
-combined_data <- bind_rows(dem_data, rep_data)
-pred_summary_dem <- new_data %>%
-  mutate(
-    pred_mean_dem = colMeans(predict_counts_dem),
-    pred_lower_dem = apply(predict_counts_dem, 2, quantile, probs = 0.025),
-    pred_upper_dem = apply(predict_counts_dem, 2, quantile, probs = 0.975)
-  )
-
-# Summarize predictions for REP
-pred_summary_rep <- new_data %>%
-  mutate(
-    pred_mean_rep = colMeans(predict_counts_rep),
-    pred_lower_rep = apply(predict_counts_rep, 2, quantile, probs = 0.025),
-    pred_upper_rep = apply(predict_counts_rep, 2, quantile, probs = 0.975)
-  )
-
-# Combine both summaries
-combined_summary <- pred_summary_dem %>%
-  select(days_since_start, pred_mean_dem, pred_lower_dem, pred_upper_dem) %>%
-  left_join(
-    pred_summary_rep %>% 
-      select(days_since_start, pred_mean_rep, pred_lower_rep, pred_upper_rep),
-    by = "days_since_start"
-  )
-# Extract the last day data for annotations
-last_day_data <- combined_summary %>%
-  filter(days_since_start == max(days_since_start))
-```
-
-```{r}
-#| echo: false
-#| label: fig-baypredict
-#| fig-cap: "Predicted number of supporters for DEM and REP party"
-#| warning: false
-
-ggplot(combined_summary, aes(x = days_since_start)) +
-  geom_line(aes(y = pred_mean_dem, color = "DEM"), size = 1) +
-  geom_ribbon(aes(ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"), alpha = 0.2) +
-  geom_line(aes(y = pred_mean_rep, color = "REP"), size = 1) +
-  geom_ribbon(aes(ymin = pred_lower_rep, ymax = pred_upper_rep, fill = "REP"), alpha = 0.2) +
-  scale_color_manual(values = c("DEM" = "blue", "REP" = "red")) +
-  scale_fill_manual(values = c("DEM" = "blue", "REP" = "red")) +
-  geom_text(
-    data = last_day_data,
-    aes(
-      x = days_since_start,
-      y = pred_mean_dem,
-      label = paste("DEM:", round(pred_mean_dem))
-    ),
-    color = "blue",
-    vjust = -1,
-    hjust = 1
-  ) +
-  geom_text(
-    data = last_day_data,
-    aes(
-      x = days_since_start,
-      y = pred_mean_rep,
-      label = paste("REP:", round(pred_mean_rep))
-    ),
-    color = "red",
-    vjust = -1,
-    hjust = 1
-  ) +
-  labs(
-    x = "Days Since Start",
-    y = "Predicted Number of Supporters",
-    color = "Party",
-    fill = "Party"
-  ) +
-  theme_minimal()
-```
-
-The prediction results of the Bayesian model closely align with those of the linear model, with both showing the mean predicted number of supporters for each party.
-Notably, the Democratic Party takes the lead around day 210 (late July).
-By early November, the predictions indicate that the Democratic Party has an average of 601 supporters, compared to the Republican Party's 578.
-This suggests a slight advantage for the Democrats as the final prediction period concludes.
-
-## Spline Fit model set-up
-
-We first run the model in R [@citeR] using the `rstanarm` package of [@rstanarm], and use priors $\text{Normal}(0, 5)$ to allow for more flexibility of predictors effects with “autoscale = TRUE”.
-Then we use the same data frame that used for predict for Bayesian model [@tbl-newdata], the result is shown as [@sec-splresult]
 
@@
 #| echo: false
+#| label: fig-splpredict
+#| fig-cap: "Poll Percentage over Time with Spline Fit for DEM and REP party"
 #| warning: false
-#| message: false
 
@@
 
-# Extract the last day data for annotations
-last_day_data <- pred_summary %>%
-  filter(days_since_start == max(days_since_start))
-```
-
-```{r}
-#| echo: false
-#| label: fig-splpredict
-#| fig-cap: "Predicted support percentage over Time with Spline Fit for DEM and REP party"
-#| warning: false
-
-ggplot(pred_summary, aes(x = days_since_start)) +
-  geom_line(aes(y = pred_mean_dem, color = "DEM"), size = 1, data = subset(pred_summary, party == "DEM")) +
-  geom_ribbon(aes(ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"), alpha = 0.2, data = subset(pred_summary, party == "DEM")) +
-  geom_line(aes(y = pred_mean_rep, color = "REP"), size = 1, data = subset(pred_summary, party == "REP")) +
-  geom_ribbon(aes(ymin = pred_lower_rep, ymax = pred_upper_rep, fill = "REP"), alpha = 0.2, data = subset(pred_summary, party == "REP")) +
-  scale_color_manual(values = c("DEM" = "blue", "REP" = "red")) +
-  scale_fill_manual(values = c("DEM" = "blue", "REP" = "red")) +
-    geom_text(
-    data = last_day_data,
-    aes(
-      x = days_since_start,
-      y = pred_mean_dem,
-      label = paste("DEM:", round(pred_mean_dem, 2))
-    ),
-    color = "blue",
-    vjust = -1,
-    hjust = 1
+# Plot the spline fit
+ggplot(combined_data, aes(x = days_since_start, y = pct, color = pollster)) +
+  # DEM predictions
+  geom_line(
+    data = pred_summary %>% filter(party == "DEM"),
+    aes(x = days_since_start, y = pred_mean_dem, color = "DEM"),
+    inherit.aes = FALSE
   ) +
-  geom_text(
-    data = last_day_data,
-    aes(
-      x = days_since_start,
-      y = pred_mean_rep,
-      label = paste("REP:", round(pred_mean_rep, 2))
-    ),
-    color = "red",
-    vjust = -1,
-    hjust = 1
+  geom_ribbon(
+    data = pred_summary %>% filter(party == "DEM"),
+    aes(x = days_since_start, ymin = pred_lower_dem, ymax = pred_upper_dem, fill = "DEM"),
+    alpha = 0.2,
+    inherit.aes = FALSE
   ) +
+  # REP predictions
+  geom_line(
+    data = pred_summary %>% filter(party == "REP"),
+    aes(x = days_since_start, y = pred_mean_rep, color = "REP"),  # Move color inside aes()
+    inherit.aes = FALSE
+  ) +
+  geom_ribbon(
+    data = pred_summary %>% filter(party == "REP"),
+    aes(x = days_since_start, ymin = pred_lower_rep, ymax = pred_upper_rep, fill = "REP"),  # Move fill inside aes()
+    alpha = 0.2,
+    inherit.aes = FALSE
+  ) +
+  # Use scale_color_manual and scale_fill_manual for colors and fills
+  scale_color_manual(values = c("DEM" = "blue", "REP" = "red"), 
+                     name = "Party",
+                     labels = c("DEM (Blue)", "REP (Red)")) +
+  scale_fill_manual(values = c("DEM" = "blue", "REP" = "red"), 
+                    name = "Party",
+                    labels = c("DEM (Blue)", "REP (Red)")) +
   labs(
-    x = "Days Since Start",
-    y = "Predicted Number of Supporters",
-    color = "Party",
-    fill = "Party"
+    x = "Days since 2024",
+    y = "Percentage",
   ) +
@@
 The prediction graph [@fig-splpredict] contains prediction of spline fit model for two parties, blue line represents the DEM party's predicted percentage of support, and the red line represents the REP party's predicted percentage of support over time (measured in days since the start of 2024).
-We can tell that the predicted support of DEM party is taking the lead by around 2 percent until the last day of the dataset.
+We can tell that the predicted support of DEM party is taking the lead by aroud 2 percent until the last day of the dataset.
 
@@
 #| label: fig-ppcheck
-#| fig-cap: "Posterior prediction check for Bayesian models"
-#| layout-ncol: 2
-#| fig-subcap: ["Democratic Model", "Republican Model"]
+#| fig-cap: "Posterior prediction check for Bayesian model"
 
-pp_check(bayesian_model_dem)
-pp_check(bayesian_model_rep)
+pp_check(Bay_model)
+
 ```

### 2024-11-04 scripts/03-clean_data.R
unused part in clean data deleted

@@ -36,27 +36,5 @@ cleaned_data <-
   mutate(
     state = if_else(is.na(state), "National", state))
 
-
-#### Prepare dataset for Harris only ####
-# Read in the data and clean variable names
-data <- read_csv("data/01-raw_data/raw_data.csv") |>
-  janitor::clean_names()
-
-# Filter data to Harris estimates based on high-quality polls after she declared
-just_harris_high_quality <- data |>
-  filter(
-    candidate_name == "Kamala Harris",
-    numeric_grade >= mean(!is.na(numeric_grade))
-  ) |>
-  mutate(
-    state = if_else(is.na(state), "National", state),
-    end_date = mdy(end_date)
-  ) |>
-  filter(end_date >= as.Date("2024-07-21")) |> 
-  mutate(
-    num_harris = round((pct / 100) * sample_size, 0) 
-  )
-
 #### Save data ####
 write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")
-write_csv(just_harris_high_quality, "data/02-analysis_data/just_harris_high_quality.csv")

### 2024-11-04 paper/paper.qmd
introduction and abstract updated

@@
 abstract: "we use FiveThirtyEight’s 2024 U.S.
-presidential poll-of-polls data to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
-By focusing on key reliability metrics—such as sample size, numeric scores, poll scores, and state-level factors—we fit the data into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
+presidential poll-of-polls data from [date] to [date] to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
+By focusing on key reliability metrics, such as sample size, numeric scores, poll scores, and state-level factors.
+We fit the data into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
 Our findings reveal a strong lead for Harris in both the linear regression and Bayesian models, with a more modest lead in the spline-fit Gaussian model.
@@
 # Introduction
+
+Nowadays, U.S.
+presidential elections have been marked by intense public interest and polarization.
+The race between Donald Trump and Kamala Harris, both of them are having a significant leads over other candidates, is currently the most closely watched contest.
+Polling data is now playing a very important role in showing the public perceptions and support rate to each candidates.
+Analyzing polling data with statistical models, we are able to gain a valuable insight into candidate support rate across different states and parties.
+These insights could help the public to predict the shifts in national, economic and social directions.
+Since the stances hold by Trump and Harris on key issues like healthcare, immigration, and the international relations are in distinct positions, analyzing on polling data to predict the winner of the 2024 U.S.
+presidential election is been crucial.
+This paper is contributing to this understanding using the latest polling data and statistical models to capture the trends in support rates of these two candidates.
+
+In this paper, polling data of 2024 U.S.
+presidential election from FiveThirtyEight [@polls], focusing specifically on the support rate of the leading candidates, Kamala Harris and Donald Trump, is analyzed.
+The polling data from FiveThirtyEight [@polls] aggregates polls from various sources, thus, to ensure the reliability on the forecast result and analysis, we reduced the variables of the data to only retain variables that reflect polling reliability, such as sample size, numeric scores and poll scores.
+Additionally, the state-level data is kept as a factor influencing support rates, acknowledging the geographic variations in voter preferences.
+
+The primary estimand in this study is the expected percentage of support (pct) for a U.S.
+presidential candidate, given polling data and relevant predictor variables like rating of the pollster and sample size of the poll.
+
+The analysis employs three statistical models: linear regression model, Bayesian model and spline-fit Gaussian model.
+The findings indicates that Kamala Harris maintains a strong lead in both linear regression model and Bayesian model, while her leads becomes modest when comparing the trend of support rate of Donald Trump in spline-fit Gaussian model.
+Comparing to Donald Trump, Kamala Harris has a [pct] lead in linear regression model, [pct] in Bayesian model and only [pct] lead in spline-fit Gaussian model.
+
+These results are significant as they provide insights into the evolving dynamics of the election and how different statistical models and approaches could yield varying interpretations of candidate support rate.
+This paper is giving public an idea of likely winner of the election that will have the potential implications on futrure political strategies and the overall direction of U.S.
+policy.
+
+The remainder of this paper is structured as follows.
+@sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
+@sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
+@sec-predict presents our *** predictions based on the model outputs.
+@sec-discuss discusses the implications of our findings and suggests directions for future research.
+Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.
 

### 2024-11-04 paper/paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
 date-format: long
-abstract: "This paper presents a multiple linear regression model for forecasting the outcome of the 2024 U.S.
-Presidential Election.
-Using nationwide polling data, we predict the percentage of support for U.S.
-presidential candidates based on key factors such as sample size, pollster quality, and timing.
-Our model aggregates these predictions to simulate potential election outcomes.
-The results reveal how specific poll attributes affect the accuracy of forecasts, offering valuable insights for predicting the likelihood of various electoral scenarios."
+abstract: "we use FiveThirtyEight’s 2024 U.S.
+presidential poll-of-polls data from [date] to [date] to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
+By focusing on key reliability metrics, such as sample size, numeric scores, poll scores, and state-level factors.
+We fit the data into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
+Our findings reveal a strong lead for Harris in both the linear regression and Bayesian models, with a more modest lead in the spline-fit Gaussian model.
+These results offer insight into the potential trajectory of the election, helping the public anticipate the policy directions that might follow a Harris administration."
 number-sections: true

### 2024-11-04 paper/paper.qmd
prose checked

@@
 date-format: long
-abstract: "we use FiveThirtyEight’s 2024 U.S.
+abstract: "we utilize FiveThirtyEight’s 2024 U.S.
 presidential poll-of-polls data from [date] to [date] to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
-By focusing on key reliability metrics, such as sample size, numeric scores, poll scores, and state-level factors.
-We fit the data into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
-Our findings reveal a strong lead for Harris in both the linear regression and Bayesian models, with a more modest lead in the spline-fit Gaussian model.
-These results offer insight into the potential trajectory of the election, helping the public anticipate the policy directions that might follow a Harris administration."
+By focusing on variables shows the reliability of a pollster, such as sample size, numeric scores, poll scores, and state-level factors, the data is fit into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
+The results of the analysis indicated a strong lead for Harris in both the linear regression and Bayesian models, with a more modest lead in the spline-fit Gaussian model.
+This paper predicts the potential trajectory of the election, helping the public anticipate the policy directions that might follow a Harris administration."
 number-sections: true
@@
 Polling data is now playing a very important role in showing the public perceptions and support rate to each candidates.
-Analyzing polling data with statistical models, we are able to gain a valuable insight into candidate support rate across different states and parties.
-These insights could help the public to predict the shifts in national, economic and social directions.
-Since the stances hold by Trump and Harris on key issues like healthcare, immigration, and the international relations are in distinct positions, analyzing on polling data to predict the winner of the 2024 U.S.
-presidential election is been crucial.
+Analyzing polling data with statistical models, we are able to have a statistical approach into candidate support rate across different states and parties.
+Analysis on polling data could help the public to predict the shifts in national, economic and social directions.
+Since the stances hold by Trump and Harris on issues like healthcare, immigration, and the international relations are in distinct positions, analyzing on polling data to predict the winner of the 2024 U.S.
+presidential election is very important to the public.
 This paper is contributing to this understanding using the latest polling data and statistical models to capture the trends in support rates of these two candidates.
@@
 
-These results are significant as they provide insights into the evolving dynamics of the election and how different statistical models and approaches could yield varying interpretations of candidate support rate.
-This paper is giving public an idea of likely winner of the election that will have the potential implications on futrure political strategies and the overall direction of U.S.
+These results are significant since they show the evolving dynamics of the election and how different statistical models and approaches could yield varying interpretations of candidate support rate.
+This paper is giving public an idea of who is the likely winner of the election that will have the potential implications on future political strategies and the overall direction of U.S.
 policy.

### 2024-11-04 paper/paper.qmd
update predict result in paper

@@
 The findings indicates that Kamala Harris maintains a strong lead in both linear regression model and Bayesian model, while her leads becomes modest when comparing the trend of support rate of Donald Trump in spline-fit Gaussian model.
-Comparing to Donald Trump, Kamala Harris has a [pct] lead in linear regression model, [pct] in Bayesian model and only [pct] lead in spline-fit Gaussian model.
+Comparing to Donald Trump, Kamala Harris has a 2.6 percent lead in linear regression model, 23 more supporter out of 1200 in Bayesian model and only 1.06 percent lead in spline-fit Gaussian model.
 
@@
 Based on the summary of models for DEM and REP [@sec-model-sigcheck], we are not sure if those predictors that measure the quality of polls are significant, so we keep them for now in the Bayesian model.
-The prediction of MLR models is shown in [@fig-mlrpredict], in which we can see the blue line represents the Democratic party has a steeper slope and starts to take the lead between day 100 and day 200.
+The prediction of MLR models is shown in [@fig-mlrpredict], in which we can see the blue line represents the Democratic party has a steeper slope and starts to take the lead between day 100 and day 200, and the average support percentage on the last day of data shows that Democratic (Harris) has a slight advantage of 2.6 percent.
 
@@
 
+# Take average of pct for each day
+combined_data_avg <- combined_data %>%
+  group_by(days_since_start) %>%
+  summarize(
+    avg_fitted_pct_dem = mean(fitted_pct_dem, na.rm = TRUE),
+    avg_fitted_pct_rep = mean(fitted_pct_rep, na.rm = TRUE)
+  )
+
+# Extract the last day data for annotations
+last_day_data <- combined_data_avg %>%
+  filter(days_since_start == max(days_since_start))
+
 # Plot with colors based on the party
-ggplot(combined_data, aes(x = days_since_start)) +
-  geom_line(aes(y = fitted_pct_dem, color = "DEM")) +
-  geom_line(aes(y = fitted_pct_rep, color = "REP")) +
+ggplot(combined_data_avg, aes(x = days_since_start)) +
+  geom_line(aes(y = avg_fitted_pct_dem, color = "DEM")) +
+  geom_line(aes(y = avg_fitted_pct_rep, color = "REP")) +
   scale_color_manual(values = c("DEM" = "blue", "REP" = "red"), 
@@
                      labels = c("DEM (Blue)", "REP (Red)")) +
+  geom_text(
+    data = last_day_data,
+    aes(
+      x = days_since_start,
+      y = avg_fitted_pct_dem,
+      label = paste("DEM:", round(avg_fitted_pct_dem, 1))
+    ),
+    color = "blue",
+    vjust = -0.5,
+    hjust = 1
+  ) +
+  geom_text(
+    data = last_day_data,
+    aes(
+      x = days_since_start,
+      y = avg_fitted_pct_rep,
+      label = paste("REP:", round(avg_fitted_pct_rep, 1))
+    ),
+    color = "red",
+    vjust = -0.5,
+    hjust = 1
+  ) +
   theme_classic() +
@@
 
+## Second discussion point
+
+Please don't use these as sub-heading labels - change them to be what your point actually is.
+
+## Third discussion point
+
 ## Weaknesses and next steps
@@
 This limitation restricts the accuracy of state-by-state analysis.
-With more reliable and updated data, using a state-based approach would enhance the model’s precision and provide a more comprehensive prediction of the election outcome.
+With more reliable and updated data, using a state-based approach would enhance the model’s precision and provide a more comprehensive prediction of the election outcome, especially for discovering the public opinion in battleground states.
 Furthermore modeling Biden and Harris together under one Democratic model introduces potential inaccuracies, especially influencing the slope of predictions in both linear and Bayesian models, making the observed upward trend less reliable.

### 2024-11-04 scripts/04-test_cleaned_data.R
cleaned data test updated

@@ -11,8 +11,9 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(testthat)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/cleaned_data.csv")
 
 # Test if the data was successfully loaded
 if (exists("analysis_data")) {
@@ -24,23 +25,17 @@ if (exists("analysis_data")) {
 
 #### Test data ####
 
+# Test that the dataset has 10 columns
+test_that("dataset has 10 columns", {
+  expect_equal(ncol(analysis_data), 10)
+})
 
+# Test that the 'state' column is character type
+test_that("'state' is character", {
+  expect_type(analysis_data$state, "character")
+})
 
-# Check if the dataset has 4 columns
-if (ncol(analysis_data) == 4) {
-  message("Test Passed: The dataset has 4 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 4 columns.")
-}
-
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
-} else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
-}
-
-# Check if the 'state' column contains only valid United state names
+# Test that 'state' column contains only valid United state names
 valid_states <- c(
   "Alabama",
   "Alaska",
@@ -91,41 +86,35 @@ valid_states <- c(
   "Washington",
   "West Virginia",
   "Wisconsin",
-  "Wyoming"
+  "Wyoming",
+  "National"
 )
+test_that("'state' contains valid United State state names", {
+  expect_true(all(analysis_data$state %in% valid_states))
+})
 
-if (all(analysis_data$state %in% valid_states)) {
-  message("Test Passed: The 'state' column contains only valid United states state names.")
-} else {
-  stop("Test Failed: The 'state' column contains invalid state names.")
-}
+# Test that the 'pollster' column is character type
+test_that("'pollster' is character", {
+  expect_type(analysis_data$pollster, "character")
+})
 
-# Check if the 'party' column contains only valid party names
-valid_parties <- c("Republican", "Democratic", "Greens", "Libertarians", "Constitution", "Natural Law")
+# Test that the 'pollster' column is character type
+test_that("'pollster' is character", {
+  expect_type(analysis_data$pollster, "character")
+})
 
-if (all(analysis_data$party %in% valid_parties)) {
[…]

### 2024-11-04 scripts/04-test_cleaned_data.R
wrong name of state fixed

@@ -51,7 +51,7 @@ valid_states <- c(
   "Idaho",
   "Illinois",
   "Indiana",
-  "Lowa",
+  "Iowa",
   "Kansas",
   "Kentucky",
   "Louisiana",
@@ -87,7 +87,12 @@ valid_states <- c(
   "West Virginia",
   "Wisconsin",
   "Wyoming",
-  "National"
+  "National",
+  "Maine CD-2",
+  "Maine CD-1",
+  "Nebraska CD-1",
+  "Nebraska CD-2",
+  "Nebraska CD-3"
 )
 test_that("'state' contains valid United State state names", {
   expect_true(all(analysis_data$state %in% valid_states))

### 2024-11-04 paper/paper.qmd
add some reference

@@
 date-format: long
-abstract: "we utilize FiveThirtyEight’s 2024 U.S.
-presidential poll-of-polls data from [date] to [date] to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
+abstract: "We utilize FiveThirtyEight’s 2024 U.S.
+presidential poll-of-polls data in 2024 to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
 By focusing on variables shows the reliability of a pollster, such as sample size, numeric scores, poll scores, and state-level factors, the data is fit into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
@@
 library(splines)
+library(broom)
+library(knitr)
 ```
@@
 Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
-Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse]Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
+Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse], packages **broom** [@broom] and **knitr** [@knitr] were used to generate clean summary and table in this paper.
+Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
 We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
@@
 
-plot(MLR_models[['DEM']], which = 1, main = "Residuals vs Fitted for DEM Party")
+plot(MLR_models[['DEM']], which = 1)
 ```
@@
 
-plot(MLR_models[['REP']], which = 1, main = "Residuals vs Fitted for REP Party")
+plot(MLR_models[['REP']], which = 1)
 ```
@@
 
-summary(MLR_models[['DEM']])
+model_summary <- tidy(MLR_models[['DEM']])
+kable(model_summary)
 ```
@@
 
-summary(MLR_models[['REP']])
+model_summary <- tidy(MLR_models[['REP ']])
+kable(model_summary)
 ```
@@
 
-## Diagnostics
-
-<!-- @fig-stanareyouokay-1 is a trace plot. It shows... This suggests... -->
-
-<!-- @fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests... -->
-
 ```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-# plot(first_model, "trace")
-# 
-# plot(first_model, "rhat")
+plot(bayesian_model_dem, pars = "(Intercept)", prob = 0.95)
 ```

### 2024-11-04 scripts/06-model_data.R

@@ -48,27 +48,8 @@ for (party in unique_parties) {
   models[[party]] <- model
 }
 
-for (party in names(models)) {
-  # Extract the model for the current party
-  current_model <- models[[party]]
 
-  # Create a Residuals vs Fitted plot for the current model
-  plot(current_model, which = 1, main = paste("Residuals vs Fitted for", party, "Party"))
-
-  # Optionally, pause between plots to view them
-  readline(prompt = "Press [Enter] to continue to the next plot...")
-}
-
-# choose a party to check summary
-model_summary <- summary(models[['DEM']])
-# Access the R-squared and Adjusted R-squared values
-r_squared <- model_summary$r.squared
-adjusted_r_squared <- model_summary$adj.r.squared
-
-# Print the model summary
-print(model_summary)
-
-# Need number not percent for some models
+# Need number not percent for Bayesian models
 nationwide_data <- nationwide_data %>% mutate(
   num_party = round((pct / 100) * sample_size, 0)
 )
@@ -98,18 +79,6 @@ dem_data <- dem_data |>
 rep_data <- rep_data |>
   mutate(fitted_pct_rep = predict(models[['REP']]))
 
-combined_data <- bind_rows(dem_data, rep_data)
-ggplot(combined_data, aes(x = days_since_start)) +
-#  geom_point(aes(y = pct), color = "black") +
-  geom_line(aes(y = fitted_pct_dem), data = subset(combined_data, party == "DEM"), color = "blue") +
-  geom_line(aes(y = fitted_pct_rep), data = subset(combined_data, party == "REP"), color = "red") +
-  theme_classic() +
-  labs(y = "Percent", x = "Date", title = "Linear Model: 
-       pct ~ log(sample_size) + pollscore + numeric_grade + transparency_score + 
-       days_since_start")
-
-modelsummary(models = models)
-
 # Bayesian model section 
 model_formula_dem <- cbind(num_party, sample_size - num_party) ~ log(sample_size) + days_since_start + 
    pollscore + numeric_grade + transparency_score + (1 | pollster)
@@ -138,9 +107,6 @@ bayesian_model_rep <- stan_glmer(
   adapt_delta = 0.95
 )
 
-# Posterior predictive checks
-pp_check(bayesian_model_dem)
-
 plot(bayesian_model_dem, pars = "(Intercept)", prob = 0.95)
 
 spline_model_dem <- stan_glm(
[…]

### 2024-11-04 paper/paper.qmd
data overview updated

--- 2024-11-04 paper/paper.qmd
@@
   - Claire Chang
-thanks: "Code and data are available at: https://github.com/zzq20010617/2024_USelection_prediction"
+thanks: "Code and data are available at: c"
 date: today
@@
 
+### Data Tools
+
 Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models] to process the data.
@@
 We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
+
+### Cleaned Data
+
+The raw data obtained from FiveThirtyEight [@polls] initially contains 17,440 observations, which we reduced to 12,185 by only retaining polls conducted after January 1, 2024.
+As our analysis focuses specifically on Kamala Harris and Donald Trump, data collected before January 1, 2024, holds limited relevance due to the inclusion of outdated candidates.
+Reducing the data further to only include polls after Harris officially announced her candidacy would result in too few observations to support a robust model.
+Additionally, the "state" column in the dataset was adjusted by replacing all "NA" values with "National," as these "NA" entries represent polls conducted nationwide in the raw data.
+A view of the first five rows of cleaned data is shown in @tbl-cleandatarows
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-cleandatarows
+#| tbl-cap: "First Five Rows of Cleaned Data"
+#| warning: false
+
+five_rows <- as_tibble(cleaned_data)
+five_rows %>% 
+  slice(1:5)
+
+```
 

### 2024-11-04 paper/paper.qmd
accidentally deleted the repo link

@@
   - Claire Chang
-thanks: "Code and data are available at: c"
+thanks: "Code and data are available at: https://github.com/zzq20010617/2024_USelection_prediction"
 date: today

### 2024-11-04 paper/paper.qmd
some sentences deleted

@@
 Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
-Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse]Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
-We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
+Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse].
 

### 2024-11-04 paper/paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
   - Claire Chang
-thanks: "Code and data are available at: c"
+thanks: "Code and data are available at: https://github.com/zzq20010617/2024_USelection_prediction"
 date: today
 date-format: long
-abstract: "we utilize FiveThirtyEight’s 2024 U.S.
-presidential poll-of-polls data from [date] to [date] to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
+abstract: "We utilize FiveThirtyEight’s 2024 U.S.
+presidential poll-of-polls data in 2024 to analyze support trends for the leading candidates, Kamala Harris and Donald Trump.
 By focusing on variables shows the reliability of a pollster, such as sample size, numeric scores, poll scores, and state-level factors, the data is fit into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model.
@@
 
-### Data Tools
-
 Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models] to process the data.
 Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
-Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse]Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
+Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse], packages **broom** [@broom] and **knitr** [@knitr] were used to generate clean summary and table in this paper.
+Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
 We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
 
-### Cleaned Data
+## Cleaned Data
 

### 2024-11-04 paper/paper.qmd
update llm usage

@@
 
-This paper examines the development of a multiple linear regression (MLR) model to predict the percentage of support (pct) for U.S.
+<!-- This paper examines the development of a multiple linear regression (MLR) model to predict the percentage of support (pct) for U.S.
 presidential candidates based on polling data.
@@
 The goal is to provide a clearer understanding of how various poll attributes influence polling results and to derive insights that can predict election outcomes.
+-->
 
-The primary estimand in this study is the expected percentage of support (pct) for a U.S.
+<!-- The primary estimand in this study is the expected percentage of support (pct) for a U.S.
 presidential candidate, given polling data and relevant predictor variables like rating of the pollster and sample size of the poll.
+-->
 
-Our Results shows that
+<!-- Our Results shows that -->
 
-This analysis is useful in understanding how various polling factors contribute to predicting election results.
+<!-- This analysis is useful in understanding how various polling factors contribute to predicting election results.
 By modeling the relationship between polling result and factors, the study enhances the ability to forecast election outcomes based on public opinion data.
 This model provides a practical tool for researchers, political strategists, and analysts to assess the reliability of polling data and its implications for elections.
+-->
 
-This paper develops statistical models to forecast the 2024 presidential election outcome between Kamala Harris and Donald Trump.
+<!-- This paper develops statistical models to forecast the 2024 presidential election outcome between Kamala Harris and Donald Trump.
 Utilizing multiple linear regression, we predict the support percentages for each candidate across various states, incorporating essential factors such as pollster rating, sample size, and state-specific demographics.
@@
 Additionally, we analyze [@trafalgar_polling_methodology] polling methodology and propose an optimized survey and methodology aimed at improving election forecast accuracy.
+-->
 
-The remainder of this paper is structured as follows.
+<!-- The remainder of this paper is structured as follows.
 @sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
@@
 Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.
+-->
 
@@
 We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
-
-## Cleaned Data
-
-The raw data obtained from FiveThirtyEight [@polls] initially contains 17,440 observations, which we reduced to 12,185 by only retaining polls conducted after January 1, 2024.
-As our analysis focuses specifically on Kamala Harris and Donald Trump, data collected before January 1, 2024, holds limited relevance due to the inclusion of outdated candidates.
-Reducing the data further to only include polls after Harris officially announced her candidacy would result in too few observations to support a robust model.
-Additionally, the "state" column in the dataset was adjusted by replacing all "NA" values with "National," as these "NA" entries represent polls conducted nationwide in the raw data.
-A view of the first five rows of cleaned data is shown in @tbl-cleandatarows
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-cleandatarows
-#| tbl-cap: "First Five Rows of Cleaned Data"
-#| warning: false
-
-five_rows <- as_tibble(cleaned_data)
-five_rows %>% 
-  slice(1:5)
-
-```
 

### 2024-11-04 paper/paper.qmd
measurement updated

--- 2024-11-04 paper/paper.qmd
@@
 library(knitr)
+library(stringr)
 ```
@@
 five_rows <- as_tibble(cleaned_data)
+five_rows <- five_rows %>%
+  rename_with(~ str_replace_all(., "_", " "))
 five_rows %>% 
@@
 
-The dataset is about public opinions about U.S. presidential candidates, which are captured through polling.
+Polling data is collected by different pollsters using various survey methods reflect public opinion on the 2024 U.S.
+presidential election.
+Information of individuals collected by pollsters is being separate to different groups based on age, gender, race, occupation and geographic to reflect a more accurate supporting trend of each candidates involved in the presidential elections.
+From FiveThirtyEight, poll-of-polls dataset included the pollscore and numeric grade of the pollster which influence the reliability of the pollster on the collection of support rate of each candidates.
+Polling methodology is also various among pollsters, such as online panels, app panels, online Ad and phone interviews.
+Different methodology are having its own target audiences which the data collected can represent the opinion of specific age group.
 

### 2024-11-04 paper/paper.qmd
modify table, add parquet for analyse data, modify readme

--- 2024-11-04 paper/paper.qmd
@@
 library(knitr)
+library(kableExtra)
 ```
@@
   filter(party == "DEM")
-dem_data <- na.omit(dem_data)
 
@@
   filter(party == "REP")
-rep_data <- na.omit(rep_data)
 
@@
 
-<!-- This paper examines the development of a multiple linear regression (MLR) model to predict the percentage of support (pct) for U.S.
-presidential candidates based on polling data.
-The data includes a range of predictors, such as sample size, pollster quality factors, and time-related factors.
-The focus of the analysis is on nationwide polling data, aggregated from different pollsters, to create a robust model for forecasting election outcomes.
-The goal is to provide a clearer understanding of how various poll attributes influence polling results and to derive insights that can predict election outcomes.
--->
-
-<!-- The primary estimand in this study is the expected percentage of support (pct) for a U.S.
-presidential candidate, given polling data and relevant predictor variables like rating of the pollster and sample size of the poll.
--->
-
-<!-- Our Results shows that -->
-
-<!-- This analysis is useful in understanding how various polling factors contribute to predicting election results.
-By modeling the relationship between polling result and factors, the study enhances the ability to forecast election outcomes based on public opinion data.
-This model provides a practical tool for researchers, political strategists, and analysts to assess the reliability of polling data and its implications for elections.
--->
-
-<!-- This paper develops statistical models to forecast the 2024 presidential election outcome between Kamala Harris and Donald Trump.
-Utilizing multiple linear regression, we predict the support percentages for each candidate across various states, incorporating essential factors such as pollster rating, sample size, and state-specific demographics.
-By aggregating these state-level predictions, we simulate possible outcomes to estimate each candidate's probability of achieving the necessary 270 electoral votes.
-Additionally, we analyze [@trafalgar_polling_methodology] polling methodology and propose an optimized survey and methodology aimed at improving election forecast accuracy.
--->
-
-<!-- The remainder of this paper is structured as follows.
-@sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
-@sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
-@sec-predict presents our *** predictions based on the model outputs.
-@sec-discuss discusses the implications of our findings and suggests directions for future research.
-Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.
--->
-
 # Data {#sec-data}
@@
 
-Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models] to process the data.
+Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models], and **arrow** [@arrow] to process the data.
 Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
 Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse], packages **broom** [@broom] and **knitr** [@knitr] were used to generate clean summary and table in this paper.
-Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
-We are creating different graphs and models for the candidates from different parties in general state and combine all models and graphs together to create a new scatter plot to predict and forecasting the US president election in 2024.
+Following **Telling stories with data** [@telling_stories_with_data], we create different models for the candidates from different parties in general state and forecasting the US president election in 2024.
 
@@
 
-five_rows <- as_tibble(cleaned_data)
-five_rows %>% 
-  slice(1:5)
-
+cleaned_data_subset <- head(cleaned_data, 5)
+colnames(cleaned_data_subset) <- c("pollster"	, "sample size", "poll grade", 
+                                   "pollscore","state", "transparency score", 
+                                   "poll end date", "party", "candidate", 
+                                   "percentage")
+kable(cleaned_data_subset, digits = 2) %>%
+    kable_styling(font_size = 6) 
 ```
@@
 
-```{r}
-plot(bayesian_model_dem, pars = "(Intercept)", prob = 0.95)
-```
-
 \newpage

### 2024-11-04 paper/paper.qmd
upadate reference and predictor section

--- 2024-11-04 paper/paper.qmd
@@
 library(stringr)
+library(patchwork)
 ```
@@
 Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models] to process the data.
-Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files.
+Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files, package **stringr** [@stringr] is used to clean the column names in tables.
 Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse], packages **broom** [@broom] and **knitr** [@knitr] were used to generate clean summary and table in this paper.
+Package **patchwork** [@patchwork] is been used to combine two graphs together.
 Following **Telling stories with data** [@telling_stories_with_data], we consider using multiple linear model to predict and forecasting the result of US president election in 2024.
@@
 Our primary outcome variable is **support percentage (pct)**, which represents the percentage of respondents who support each candidate.
-This variable is modeled as the response variable in the MLR analyse.
+This variable is modeled as the response variable in the linear regression model, Bayesian model and spline-fit Gaussian model analysis.
 
@@
 Sample size is an important predictor because it influences the reliability of a poll.
-Larger sample sizes tend to reduce sampling error, providing more accurate reflections of voter sentiment.
-In our MLR model, the sample size is log-transformed to account for diminishing returns—larger polls do not necessarily offer proportionally better accuracy.
+Larger sample sizes tend to reduce sampling error, providing more accurate reflections of voter sentiment and more reliable predictions of the election.
+In our models, the sample size is log-transformed to account for diminishing returns—larger polls do not necessarily offer proportionally better accuracy.
 
-### Poll Score
+### Poll Score and Numeric Grade
 
 The poll score is a measure of the error and bias we can attribute to a pollster, negative number is better.
-
-### Numeric Grade
-
-This variable is an aggregate score of the poll based on a numeric scale.
+The numeric grade is an aggregate score of the poll based on a numeric scale.
 It serves as an indicator of the pollster's historical performance and the methodology used.
 3 is the maximum and some pollsters have no rating.
+A lower poll score and higher numeric grade of a pollster indicate the data collected by the pollster has higher reliability and the methodology used by the pollster is more consistent.
+
+```{r}
+#| label: fig-pollscore-numericgrade
+#| fig-cap: "Distribution of Poll scores and Numeric Grade of Pollsters in the 2024 U.S. Presidential Election"
+#| echo: false
+#| warning: false
+
+ps_graph <- ggplot(cleaned_data, aes(x = pollscore)) +
+  geom_bar(fill = "blue", color = "black") +
+  labs(x = "Pollscore",
+       y = "Frequency") +
+  theme_minimal()
+
+ng_graph <- ggplot(cleaned_data, aes(x = numeric_grade)) +
+  geom_bar(fill = "red", color = "black") +
+  labs(x = "Numeric Grade",
+       y = "Frequency") +
+  theme_minimal()
+
+ps_graph + ng_graph
+```
+The left side of @fig-pollscore-numericgrade presents the distribution of poll scores among all of the pollsters, most of the pollsters have negative poll scores which indicates these pollsters are reliable.
+Among these pollsters having negative poll scores, more than 2000 pollsters have poll scores lower than -1 showing higher reliability of the data collected by them than other pollsters.
+Some of the pollsters have positive poll scores, data collected by these pollsters is not reliable.
+
+The right side of @fig-pollscore-numericgrade presents the distribution of numeric grades of all pollsters, most of the pollsters have the numeric grades higher than the middle number, 1.5.
+More than 3000 polling organizations have numeric grades higher than 2.8 reflects a higher reliability of data collected by them.
 
@@
 
+```{r}
+#| label: fig-transparency
+#| fig-cap: "Distribution of Transparency Score of Pollsters in the 2024 U.S. Presidential Election"
+#| echo: false
+#| warning: false
+
+ts_graph <- ggplot(cleaned_data, aes(x = transparency_score)) +
+  geom_bar(fill = "blue", color = "black") +
+  labs(x = "Transparency Score",
+       y = "Frequency") +
+  theme_minimal()
+
+ts_graph
+```
+From the distribution graph of transparency score among all of the pollster shown in @fig-transparency, this reflects the differences in the methodologies used by different pollsters.
+Over 3,000 pollsters have a transparency score above 9, indicating that they have more open and trustworthy methodologies.
+In contrast, some pollsters have a transparency score below 5, which suggests that their methodologies are not disclose the details of their data collection processes, thereby diminishing the credibility of these polls."
+
 ### Days Since Start
@@
 This variable represents the number of days since the beginning of the election polling period.
+It is calculated as Days Since Start = number of days from the beginning of the election to the end date of the specific poll.
 It helps capture the dynamic nature of voter preferences over time, accounting for shifts in public opinion as the election date nears.
+We only keep the polls with end dates after January 1 2024 to keep the variety of the data but avoid candidates and parties that are not involved in the final race of the election to mainly analyze the support rate percentage between Donald Trump and Kamala Harris.
 

### 2024-11-04 paper/paper.qmd
add sketches

--- 2024-11-04 paper/paper.qmd
@@
 
-## Result {#sec-splresult}
+### Result {#sec-splresult}
 
@@
 
-## Second discussion point
+## Differences in Modeling Nationwide Polls and Actual Election Results
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
+In this paper, we used nationwide poll data for modeling, which contrasts significantly with using state-by-state election results.
+Nationwide polls offer general insights into aggregate support levels but do not account for the Electoral College system in U.S.
+presidential elections.
+This approach can overlook state-specific variations and the unique influence of each state's electoral votes, leading to potential bias where larger states disproportionately affect overall results.
+Conversely, state-level modeling captures the distribution of support across states, highlighting the impact of battleground states.
+To enhance prediction accuracy and align with the actual election structure, future models should integrate state-level data especially for the battleground states for more robust, applicable results.
 

### 2024-11-04 paper/paper.qmd
update render pdf

@@
 library(knitr)
+library(patchwork)
 library(kableExtra)
@@
 @sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
-@sec-predict presents our *** predictions based on the model outputs.
+Our predictions are underr section of each model.
 @sec-discuss discusses the implications of our findings and suggests directions for future research.
@@
 Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse], packages **broom** [@broom] and **knitr** [@knitr] were used to generate clean summary and table in this paper.
+Package **patchwork** [@patchwork] is been used to combine two graphs together.
 Following **Telling stories with data** [@telling_stories_with_data], we create different models for the candidates from different parties in general state and forecasting the US president election in 2024.
@@
 ```
+## Cross-Model Observations
+
+In both the MLR and Bayesian models, Harris demonstrates a significant lead over Trump.
+The MLR model projects an approximate 2.6% lead for Harris, as indicated by the slope for the Democratic (DEM) variable, which is steeper and surpasses the Republican (REP) trend line between day 100 and day 200 (see @fig-mlrpredict).
+This early crossover point indicates a shift in support dynamics, suggesting that Harris’s policies or campaign events during this period may have positively influenced public sentiment.
+The Bayesian model, which accounts for random effects among pollsters, reinforces this lead by predicting 601 supporters for Harris versus 578 for Trump towards the end of the prediction period (see @fig-baypredict).
+This stability across models suggests that Harris’s support is likely durable, despite the model-specific variations in lead magnitude.
+The Spline-Fit Gaussian model, however, presents a more conservative lead for Harris, with a smaller difference of approximately 1.06% in favor of the Democratic candidate by the end of the dataset (see @fig-splpredict).
+This model, designed to capture smoother trends, suggests that Harris’s lead, while present, may not be as robust as projected by the MLR and Bayesian models.
+This discrepancy emphasizes that different modeling approaches yield varied levels of confidence in support predictions, with the Spline-Fit model capturing subtle shifts that might indicate voter hesitancy or fluctuating preferences over time.
 

### 2024-11-04 paper/paper.qmd
minor error fixed

--- 2024-11-04 paper/paper.qmd
@@
 library(patchwork)
+library(kableExtra)
 ```
@@
 @sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
-@sec-predict presents our *** predictions based on the model outputs.
-@sec-discuss discusses the implications of our findings and suggests directions for future research.
-Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.
-
-This paper examines the development of a multiple linear regression (MLR) model to predict the percentage of support (pct) for U.S.
-presidential candidates based on polling data.
-The data includes a range of predictors, such as sample size, pollster quality factors, and time-related factors.
-The focus of the analysis is on nationwide polling data, aggregated from different pollsters, to create a robust model for forecasting election outcomes.
-The goal is to provide a clearer understanding of how various poll attributes influence polling results and to derive insights that can predict election outcomes.
-
-The primary estimand in this study is the expected percentage of support (pct) for a U.S.
-presidential candidate, given polling data and relevant predictor variables like rating of the pollster and sample size of the poll.
-
-Our Results shows that
-
-This analysis is useful in understanding how various polling factors contribute to predicting election results.
-By modeling the relationship between polling result and factors, the study enhances the ability to forecast election outcomes based on public opinion data.
-This model provides a practical tool for researchers, political strategists, and analysts to assess the reliability of polling data and its implications for elections.
-
-This paper develops statistical models to forecast the 2024 presidential election outcome between Kamala Harris and Donald Trump.
-Utilizing multiple linear regression, we predict the support percentages for each candidate across various states, incorporating essential factors such as pollster rating, sample size, and state-specific demographics.
-By aggregating these state-level predictions, we simulate possible outcomes to estimate each candidate's probability of achieving the necessary 270 electoral votes.
-Additionally, we analyze [@trafalgar_polling_methodology] polling methodology and propose an optimized survey and methodology aimed at improving election forecast accuracy.
-
-The remainder of this paper is structured as follows.
-@sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
-@sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
-@sec-predict presents our *** predictions based on the model outputs.
-@sec-discuss discusses the implications of our findings and suggests directions for future research.
+Results of each model are discussed under the specific model section @sec-discuss discusses the implications of our findings and suggests directions for future research.
 Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.
@@
 
-five_rows <- as_tibble(cleaned_data)
-five_rows <- five_rows %>%
-  rename_with(~ str_replace_all(., "_", " "))
-five_rows %>% 
-  slice(1:5)
-
+cleaned_data_subset <- head(cleaned_data, 5)
+colnames(cleaned_data_subset) <- c("pollster"	, "sample size", "poll grade", 
+                                   "pollscore","state", "transparency score", 
+                                   "poll end date", "party", "candidate", 
+                                   "percentage")
+kable(cleaned_data_subset, digits = 2) %>%
+    kable_styling(font_size = 6) 
 ```
